---
title: "Sargasso_Subisidy_Code"
output:
  word_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
---

#libraries

```{r}
library(knitr)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(ggfortify)
library(lme4)
library(lmerTest)
library(car)
library(nlme)
library(ggpubr)
library(multcompView)
library(boot)
library(Hotelling)
library(mvnTest)
library(vegan)
library(factoextra)
library(FactoMineR)
library(agricolae)
library(glmm)
library(HH)
library(ggpubr)
library(rstatix)
library(vegan)
library(effsize)
library(broom)
library(performance)
library(DHARMa)
library(glmmTMB)
library(emmeans)

```


# Pile metrics

```{r}
piles <-read.csv("Sargassum_piles.csv",header=T)
piles2<-subset(piles, Treatment== "Sargassum")
piles2$Trip<-as.factor(piles2$Trip)
piles2$Block<-as.factor(piles2$Block)

# Create the scatter plot with jittered points and add a trendline
ggplot(piles2, aes(x = Trip, y = Ellipsoid_prc_D, color = Location)) +
  geom_boxplot() +
  geom_smooth(method = "lm", se = FALSE)
  

#Potential model:  repeated measures ANOVA (as in Spiller et al 2010)
#Pile volume as the response, site as the main between-subjects factor, sampling date as the repeated within-subject 
#factor, and starting volume as a random factor.

# Perform the repeated measures ANOVA with original_vol as a covariate
piles_rm_anova <- aov(ellipsoid_vol ~ Location * Trip + Error(Block / Trip) + (1|OG_ellipsoid_Vol), data = piles2)

summary(piles_rm_anova)

```

### pile figures
```{r}
piles <-read.csv("Sargassum_piles.csv",header=T)
piles2<-subset(piles, Treatment== "Sargassum")
piles2$Trip<-as.factor(piles2$Trip)
piles2$Block<-as.factor(piles2$Block)
# Calculate the means by month and location

mean_data <- piles2 %>%
  group_by(Trip, Location) %>%
  summarize(mean_ellipsoid_vol = mean(Ellipsoid_prc_D))


# Create the scatter plot with mean points connected by a line
vol_scatterplot <- ggplot(data = piles2, aes(x = Trip, y = Ellipsoid_prc_D, fill = Location)) +
  geom_point(size = 2, shape = 21, aes(fill = Location)) +
  geom_line(data = mean_data, aes(x = Trip, y = mean_ellipsoid_vol, group = Location), color = "gray")

# Customize the plot appearance
vol_x_time<- vol_scatterplot + 
  labs(x = "", y = "Pile volume (%)") +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_manual(values = c("black", "white")) +
   scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar.23", "4" = "Aug.23")
  ) +
  theme_minimal() + 
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"))
vol_x_time



```


#Decomposer experiment

```{r}

decomp<-read.csv("mesh_bag_exp.csv",header = T )
decomp$Rep_no.<-as.factor(decomp$Rep_no.)
decomp$Trip<-as.factor(decomp$Trip)
decomp$prc_Delta<-as.numeric(decomp$prc_Delta)

#quick view counts of percent
ggplot(decomp) +
 aes(x = Trip, fill = Treatment, weight = prc_means) +
 geom_bar(position = "dodge") +
 scale_fill_hue(direction = 1) +
 theme_minimal() +
 facet_grid(vars(Site), vars())


#now lets look at mean values and SE
#subset out trip 1 since its only going to show initial mass which is 100%
decomp %>%
  filter(!(Trip %in% "1")) %>%
  filter(!(ratio %in% "")) %>%
  ggplot() +
  scale_fill_manual(name = "Mesh size", 
                    values = c("large" = "grey25", "small" = "dark gray")) +
  aes(x = Site, fill = Treatment, weight = prc_means) +
  geom_bar(position = "dodge") +
  geom_errorbar(aes(ymin = prc_means - prc_SE, ymax = prc_means + prc_SE), 
                position = position_dodge(width = 0.9), 
                width = 0.25) +  
  labs(y = "Decomposition (% of initial dry biomass)") +
  theme_minimal() +
  facet_grid(Trip ~ ., labeller = labeller(Trip = function(x) ifelse(x == "2", "Nov. 22", ifelse(x == "3", "Mar. 23", x)))) 



```


#Plant Communities 
## Interior survey

```{r}
#upload and subset data for interior surveys 
#import dataset
interiorsurvey<-read.csv("interior_survey.csv", header = TRUE)

is.factor(interiorsurvey$Treatment)
interiorsurvey$Treatment<-as.factor(interiorsurvey$Treatment)
is.factor(interiorsurvey$Block)
interiorsurvey$Block<-as.factor(interiorsurvey$Block)
interiorsurvey$Trip<-as.factor(interiorsurvey$Trip)


###########subset out the dunes
interior_sub_dune<-subset(interiorsurvey, Site == "Dunes")

#subset trip 1
int_sub_dune_1<-subset(interior_sub_dune, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
intdune_df_sargassum <- int_sub_dune_1[int_sub_dune_1$Treatment == "Sargassum", ]
intdune_df_control <- int_sub_dune_1[int_sub_dune_1$Treatment == "Control", ]

### effect sizes for trip 1
cohen.d(intdune_df_sargassum$bermuda,intdune_df_control$bermuda)
cohen.d(intdune_df_sargassum$other_plant_cover,intdune_df_control$other_plant_cover)


#subset trip 2
int_sub_dune_2<-subset(interior_sub_dune, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
intdune_df_sargassum2 <- int_sub_dune_2[int_sub_dune_2$Treatment == "Sargassum", ]
intdune_df_control2 <- int_sub_dune_2[int_sub_dune_2$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(intdune_df_sargassum2$bermuda,intdune_df_control2$bermuda)
cohen.d(intdune_df_sargassum2$total_plant_prc_cover,intdune_df_control2$total_plant_prc_cover)
cohen.d(intdune_df_sargassum2$bare_prc_cover,intdune_df_control2$bare_prc_cover)
cohen.d(intdune_df_sargassum2$leaf_prc_cover,intdune_df_control2$leaf_prc_cover)
cohen.d(intdune_df_sargassum2$other_plant_cover,intdune_df_control2$other_plant_cover)


#subset trip 3
int_sub_dune_3<-subset(interior_sub_dune, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
intdune_df_sargassum3 <- int_sub_dune_3[int_sub_dune_3$Treatment == "Sargassum", ]
intdune_df_control3 <- int_sub_dune_3[int_sub_dune_3$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(intdune_df_sargassum3$bermuda,intdune_df_control3$bermuda)
cohen.d(intdune_df_sargassum3$total_plant_prc_cover,intdune_df_control3$total_plant_prc_cover)
cohen.d(intdune_df_sargassum3$bare_prc_cover,intdune_df_control3$bare_prc_cover)
cohen.d(intdune_df_sargassum3$leaf_prc_cover,intdune_df_control3$leaf_prc_cover)
cohen.d(intdune_df_sargassum3$other_plant_cover,intdune_df_control3$other_plant_cover)


#subset trip 4
int_sub_dune_4<-subset(interior_sub_dune, Trip == "4")

# Subset the data for 'Sargassum' and 'Control' treatments
intdune_df_sargassum4 <- int_sub_dune_4[int_sub_dune_4$Treatment == "Sargassum", ]
intdune_df_control4 <- int_sub_dune_4[int_sub_dune_4$Treatment == "Control", ]

### effect sizes for trip 4
cohen.d(intdune_df_sargassum4$bermuda,intdune_df_control4$bermuda)
cohen.d(intdune_df_sargassum4$total_plant_prc_cover,intdune_df_control4$total_plant_prc_cover)
cohen.d(intdune_df_sargassum4$bare_prc_cover,intdune_df_control4$bare_prc_cover)
cohen.d(intdune_df_sargassum4$leaf_prc_cover,intdune_df_control4$leaf_prc_cover)
cohen.d(intdune_df_sargassum4$other_plant_cover,intdune_df_control4$other_plant_cover)


###########subset out the Jungle
interior_sub_Jungle<-subset(interiorsurvey, Site == "Jungle")

#subset trip 1
int_sub_Jungle_1<-subset(interior_sub_Jungle, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
intJungle_df_sargassum <- int_sub_Jungle_1[int_sub_Jungle_1$Treatment == "Sargassum", ]
intJungle_df_control <- int_sub_Jungle_1[int_sub_Jungle_1$Treatment == "Control", ]

### effect sizes for trip 1
cohen.d(intJungle_df_sargassum$bermuda,intJungle_df_control$bermuda)
cohen.d(intJungle_df_sargassum$total_plant_prc_cover,intJungle_df_control$total_plant_prc_cover)
cohen.d(intJungle_df_sargassum$bare_prc_cover,intJungle_df_control$bare_prc_cover)
cohen.d(intJungle_df_sargassum$leaf_prc_cover,intJungle_df_control$leaf_prc_cover)
cohen.d(intJungle_df_sargassum$other_plant_cover,intJungle_df_control$other_plant_cover)


#subset trip 2
int_sub_Jungle_2<-subset(interior_sub_Jungle, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
intJungle_df_sargassum2 <- int_sub_Jungle_2[int_sub_Jungle_2$Treatment == "Sargassum", ]
intJungle_df_control2 <- int_sub_Jungle_2[int_sub_Jungle_2$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(intJungle_df_sargassum2$bermuda,intJungle_df_control2$bermuda)
cohen.d(intJungle_df_sargassum2$total_plant_prc_cover,intJungle_df_control2$total_plant_prc_cover)
cohen.d(intJungle_df_sargassum2$bare_prc_cover,intJungle_df_control2$bare_prc_cover)
cohen.d(intJungle_df_sargassum2$leaf_prc_cover,intJungle_df_control2$leaf_prc_cover)
cohen.d(intJungle_df_sargassum2$other_plant_cover,intJungle_df_control2$other_plant_cover)

#subset trip 3
int_sub_Jungle_3<-subset(interior_sub_Jungle, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
intJungle_df_sargassum3 <- int_sub_Jungle_3[int_sub_Jungle_3$Treatment == "Sargassum", ]
intJungle_df_control3 <- int_sub_Jungle_3[int_sub_Jungle_3$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(intJungle_df_sargassum3$bermuda,intJungle_df_control3$bermuda)
cohen.d(intJungle_df_sargassum3$total_plant_prc_cover,intJungle_df_control3$total_plant_prc_cover)
cohen.d(intJungle_df_sargassum3$bare_prc_cover,intJungle_df_control3$bare_prc_cover)
cohen.d(intJungle_df_sargassum3$leaf_prc_cover,intJungle_df_control3$leaf_prc_cover)
cohen.d(intJungle_df_sargassum3$other_plant_cover,intJungle_df_control3$other_plant_cover)



#subset trip 4
int_sub_Jungle_4<-subset(interior_sub_Jungle, Trip == "4")

# Subset the data for 'Sargassum' and 'Control' treatments
intJungle_df_sargassum4 <- int_sub_Jungle_4[int_sub_Jungle_4$Treatment == "Sargassum", ]
intJungle_df_control4 <- int_sub_Jungle_4[int_sub_Jungle_4$Treatment == "Control", ]

### effect sizes for trip 4
cohen.d(intJungle_df_sargassum4$bermuda,intJungle_df_control4$bermuda)
cohen.d(intJungle_df_sargassum4$total_plant_prc_cover,intJungle_df_control4$total_plant_prc_cover)
cohen.d(intJungle_df_sargassum4$bare_prc_cover,intJungle_df_control4$bare_prc_cover)
cohen.d(intJungle_df_sargassum4$leaf_prc_cover,intJungle_df_control4$leaf_prc_cover)
cohen.d(intJungle_df_sargassum4$other_plant_cover,intJungle_df_control4$other_plant_cover)


```

###percent cover
```{r}

#import dataset
interiorsurvey<-read.csv("interior_graphs.csv", header = TRUE)

#we are going to use a dotplot to graph means and SE of percent cover for grass and other plants
#i made a long datasheet called interior_graphs.csv with effect sizes and means (of quadrats in each plot n=3 since we are not interested in between quadrat variability)

interiorsurvey2<-subset(interiorsurvey, Data == "graph")
#make sure we have factors 
interiorsurvey2$Treatment<-as.factor(interiorsurvey2$Treatment)
interiorsurvey2$Trip<-as.factor(interiorsurvey2$Trip)
interiorsurvey2$means<-as.numeric(interiorsurvey2$means)


#graph it


interior_pc_cover<- ggplot(data = interiorsurvey2,
            aes(
              x = Trip,
              y = means,
              shape = cover_type,
              color = Treatment
            )) +
  stat_summary(geom = "point", fun = mean, position = position_dodge(width = 1),
               size = 2L, stroke = 1) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 1),
               width = 0.2) +  # Adjust the width of the error bars as needed
  labs(
    x = " ",
    y = "Mean Percent Cover",
    title = "Interior Mean Percent Cover",
    shape = "Cover type"
  ) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 23")
  ) +
  facet_grid(vars(), vars(Site)) +
  theme_minimal() +
  scale_fill_manual(values = c("grey44", "black")) +
  scale_color_manual(values = c("grey44", "black")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines")
  )

interior_pc_cover


```

###effect sizes
````{r}

interioreffsize1 <-read.csv("interior_graphs.csv",header=T)
interioreffsize <-subset(interioreffsize1, Data == "eff_size")

interioreffsize<-subset(interioreffsize, !Cover == "all_plants")
interioreffsize<-subset(interioreffsize, !Cover ==  "bare" ) 
interioreffsize<-subset(interioreffsize, !Cover == "leaf_litter")
interioreffsize$Trip<-as.factor(interioreffsize$Trip)
interioreffsize$Effect_size<-as.numeric(interioreffsize$Effect_size)


int_eff_plot <- ggplot(interioreffsize, aes(x = Trip, y = Effect_size, shape = Cover)) +
  labs(y = "Effect size", title = "") +
  geom_path(aes(group = Cover), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar.23", "4" = "Aug.23")
  ) +
    scale_shape_discrete(labels = c("grass", "other plants")) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    strip.text = element_text(size = 12)
  ) +
  xlab("") +
    theme_minimal()  +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"))

int_eff_plot


```

## Perimeter Survey
```{r}
#upload and subset data for perimeter surveys 
#import dataset
perimetersurvey<-read.csv("perimeter_survey.csv", header = TRUE)

perimetersurvey$Treatment<-as.factor(perimetersurvey$Treatment)
perimetersurvey$Block<-as.factor(perimetersurvey$Block)
perimetersurvey$Trip<-as.factor(perimetersurvey$Trip)

###########subset out the dunes
perimeter_sub_dune<-subset(perimetersurvey, Site == "Dunes")

#subset trip 1
per_sub_dune_1<-subset(perimeter_sub_dune, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
perdune_df_sargassum <- per_sub_dune_1[per_sub_dune_1$Treatment == "Sargassum", ]
perdune_df_control <- per_sub_dune_1[per_sub_dune_1$Treatment == "Control", ]

### effect sizes for trip 1
cohen.d(perdune_df_sargassum$grass_prc_cover,perdune_df_control$grass_prc_cover)
cohen.d(perdune_df_sargassum$total_plant_prc_cover,perdune_df_control$total_plant_prc_cover)
cohen.d(perdune_df_sargassum$bare_prc_cover,perdune_df_control$bare_prc_cover)
cohen.d(perdune_df_sargassum$dead_prc_cover,perdune_df_control$dead_prc_cover)
cohen.d(perdune_df_sargassum$other_plant_prc_cover,perdune_df_control$other_plant_prc_cover)


#subset trip 2
per_sub_dune_2<-subset(perimeter_sub_dune, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
perdune_df_sargassum2 <- per_sub_dune_2[per_sub_dune_2$Treatment == "Sargassum", ]
perdune_df_control2 <- per_sub_dune_2[per_sub_dune_2$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(perdune_df_sargassum2$grass_prc_cover,perdune_df_control2$grass_prc_cover)
cohen.d(perdune_df_sargassum2$total_plant_prc_cover,perdune_df_control2$total_plant_prc_cover)
cohen.d(perdune_df_sargassum2$bare_prc_cover,perdune_df_control2$bare_prc_cover)
cohen.d(perdune_df_sargassum2$dead_prc_cover,perdune_df_control2$dead_prc_cover)
cohen.d(perdune_df_sargassum2$other_plant_prc_cover,perdune_df_control2$other_plant_prc_cover)


#subset trip 3
per_sub_dune_3<-subset(perimeter_sub_dune, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
perdune_df_sargassum3 <- per_sub_dune_3[per_sub_dune_3$Treatment == "Sargassum", ]
perdune_df_control3 <- per_sub_dune_3[per_sub_dune_3$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(perdune_df_sargassum3$grass_prc_cover,perdune_df_control3$grass_prc_cover)
cohen.d(perdune_df_sargassum3$total_plant_prc_cover,perdune_df_control3$total_plant_prc_cover)
cohen.d(perdune_df_sargassum3$bare_prc_cover,perdune_df_control3$bare_prc_cover)
cohen.d(perdune_df_sargassum3$dead_prc_cover,perdune_df_control3$dead_prc_cover)
cohen.d(perdune_df_sargassum3$other_plant_prc_cover,perdune_df_control3$other_plant_prc_cover)

#subset trip 4
per_sub_dune_4<-subset(perimeter_sub_dune, Trip == "4")

# Subset the data for 'Sargassum' and 'Control' treatments
perdune_df_sargassum4 <- per_sub_dune_4[per_sub_dune_4$Treatment == "Sargassum", ]
perdune_df_control4 <- per_sub_dune_4[per_sub_dune_4$Treatment == "Control", ]

### effect sizes for trip 4
cohen.d(perdune_df_sargassum4$grass_prc_cover,perdune_df_control4$grass_prc_cover)
cohen.d(perdune_df_sargassum4$total_plant_prc_cover,perdune_df_control4$total_plant_prc_cover)
cohen.d(perdune_df_sargassum4$bare_prc_cover,perdune_df_control4$bare_prc_cover)
cohen.d(perdune_df_sargassum4$dead_prc_cover,perdune_df_control4$dead_prc_cover)
cohen.d(perdune_df_sargassum$other_plant_prc_cover,perdune_df_control4$other_plant_prc_cover)

###########subset out the Jungle
perimeter_sub_Jungle<-subset(perimetersurvey, Site == "Jungle")

#subset trip 1
per_sub_Jungle_1<-subset(perimeter_sub_Jungle, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
perJungle_df_sargassum <- per_sub_Jungle_1[per_sub_Jungle_1$Treatment == "Sargassum", ]
perJungle_df_control <- per_sub_Jungle_1[per_sub_Jungle_1$Treatment == "Control", ]

### effect sizes for trip 1
cohen.d(perJungle_df_sargassum$grass_prc_cover,perJungle_df_control$grass_prc_cover)
cohen.d(perJungle_df_sargassum$total_plant_prc_cover,perJungle_df_control$total_plant_prc_cover)
cohen.d(perJungle_df_sargassum$bare_prc_cover,perJungle_df_control$bare_prc_cover)
cohen.d(perJungle_df_sargassum$dead_prc_cover,perJungle_df_control$dead_prc_cover)
cohen.d(perJungle_df_sargassum$other_plant_prc_cover,perJungle_df_control$other_plant_prc_cover)

#subset trip 2
per_sub_Jungle_2<-subset(perimeter_sub_Jungle, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
perJungle_df_sargassum2 <- per_sub_Jungle_2[per_sub_Jungle_2$Treatment == "Sargassum", ]
perJungle_df_control2 <- per_sub_Jungle_2[per_sub_Jungle_2$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(perJungle_df_sargassum2$grass_prc_cover,perJungle_df_control2$grass_prc_cover)
cohen.d(perJungle_df_sargassum2$total_plant_prc_cover,perJungle_df_control2$total_plant_prc_cover)
cohen.d(perJungle_df_sargassum2$bare_prc_cover,perJungle_df_control2$bare_prc_cover)
cohen.d(perJungle_df_sargassum2$dead_prc_cover,perJungle_df_control2$dead_prc_cover)
cohen.d(perJungle_df_sargassum2$other_plant_prc_cover,perJungle_df_control2$other_plant_prc_cover)

#subset trip 3
per_sub_Jungle_3<-subset(perimeter_sub_Jungle, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
perJungle_df_sargassum3 <- per_sub_Jungle_3[per_sub_Jungle_3$Treatment == "Sargassum", ]
perJungle_df_control3 <- per_sub_Jungle_3[per_sub_Jungle_3$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(perJungle_df_sargassum3$grass_prc_cover,perJungle_df_control3$grass_prc_cover)
cohen.d(perJungle_df_sargassum3$total_plant_prc_cover,perJungle_df_control3$total_plant_prc_cover)
cohen.d(perJungle_df_sargassum3$bare_prc_cover,perJungle_df_control3$bare_prc_cover)
cohen.d(perJungle_df_sargassum3$dead_prc_cover,perJungle_df_control3$dead_prc_cover)
cohen.d(perJungle_df_sargassum3$other_plant_prc_cover,perJungle_df_control3$other_plant_prc_cover)

#subset trip 4
per_sub_Jungle_4<-subset(perimeter_sub_Jungle, Trip == "4")

# Subset the data for 'Sargassum' and 'Control' treatments
perJungle_df_sargassum4 <- per_sub_Jungle_4[per_sub_Jungle_4$Treatment == "Sargassum", ]
perJungle_df_control4 <- per_sub_Jungle_4[per_sub_Jungle_4$Treatment == "Control", ]

### effect sizes for trip 4
cohen.d(perJungle_df_sargassum4$grass_prc_cover,perJungle_df_control4$grass_prc_cover)
cohen.d(perJungle_df_sargassum4$total_plant_prc_cover,perJungle_df_control4$total_plant_prc_cover)
cohen.d(perJungle_df_sargassum4$bare_prc_cover,perJungle_df_control4$bare_prc_cover)
cohen.d(perJungle_df_sargassum4$dead_prc_cover,perJungle_df_control4$dead_prc_cover)
cohen.d(perJungle_df_sargassum4$other_plant_prc_cover,perJungle_df_control4$other_plant_prc_cover)


```



###percent cover
```{r}

#import dataset
perimetersurvey<-read.csv("perimeter_graphs.csv", header = TRUE)

#we are going to use a dotplot to graph means and SE of percent cover for grass and other plants
#i made a long datasheet called perimeter_graphs.csv with effect sizes and means (of quadrats in each plot n=3 since we are not interested in between quadrat variability)

perimetersurvey2<-subset(perimetersurvey, Data == "graph")
#make sure we have factors 
perimetersurvey2$Treatment<-as.factor(perimetersurvey2$Treatment)
perimetersurvey2$Trip<-as.factor(perimetersurvey2$Trip)
perimetersurvey2$means<-as.numeric(perimetersurvey2$means)


#graph it


perimeter_pc_cover<- ggplot(data = perimetersurvey2,
            aes(
              x = Trip,
              y = means,
              shape = cover_type,
              color = Treatment
            )) +
  stat_summary(geom = "point", fun = mean, position = position_dodge(width = 1),
               size = 2L, stroke = 1) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 1),
               width = 0.2) +  # Adjust the width of the error bars as needed
  labs(
    x = " ",
    y = "Mean Percent Cover",
    title = "perimeter Mean Percent Cover",
    shape = "Cover type"
  ) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 23")
  ) +
  facet_grid(vars(), vars(Site)) +
  theme_minimal() +
  scale_fill_manual(values = c("grey44", "black")) +
  scale_color_manual(values = c("grey44", "black")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines")
  )

perimeter_pc_cover


```


###effect sizes
````{r}

perimetereffsize1 <-read.csv("perimeter_graphs.csv",header=T)
perimetereffsize <-subset(perimetereffsize1, Data == "eff_size")

perimetereffsize<-subset(perimetereffsize, !Cover == "all_plants")
perimetereffsize<-subset(perimetereffsize, !Cover ==  "dead" ) 
perimetereffsize<-subset(perimetereffsize, !Cover ==  "bare" ) 
perimetereffsize$Trip<-as.factor(perimetereffsize$Trip)
perimetereffsize$Effect_size<-as.numeric(perimetereffsize$Effect_size)


per_eff_plot <- ggplot(perimetereffsize, aes(x = Trip, y = Effect_size, shape = Cover)) +
  labs(y = "Effect size", title = "") +
  geom_path(aes(group = Cover), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar.23", "4" = "Aug.23")
  ) +
    scale_shape_discrete(labels = c("grass", "other plants")) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    strip.text = element_text(size = 12)
  ) +
  xlab("") +
    theme_minimal()  +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"))

per_eff_plot


```



###Plant community nMDS 

```{r}
#upload and subset data for interior surveys 
#import dataset
plants<-read.csv("plantcommunity_NMDS.csv", header = TRUE)

#subset out only the "random" and "interior"  since those are comparable (3 quads each pile interior)
noladder<-subset(plants, !Survey == "Ladder" )
interiorsurvey<-subset(noladder, !Survey == "HiLo")
is.factor(interiorsurvey$Treatment)
interiorsurvey$Treatment<-as.factor(interiorsurvey$Treatment)
is.factor(interiorsurvey$Block)
interiorsurvey$Block<-as.factor(interiorsurvey$Block)
#compare communities differences using NMDS
#use only the data we are interested in
##plants first
nm <- (interiorsurvey[,c(8:60)])
#make matrix
matrix = as.matrix(nm)
nmds_result = metaMDS(matrix, distance = "bray")
plot(nmds_result)


# Create a new data frame with NMDS coordinates, grouping, and additional descriptor information
nmds_data <- data.frame(nmds_result$points, Site = interiorsurvey$Site, Treatment = interiorsurvey$Treatment)


plot1 <- ggplot(nmds_data, aes(x = MDS1, y = MDS2)) + 
    geom_point(size = 3, aes( shape = Site, colour = Treatment))+ 
    theme(axis.text.y = element_text(colour = "black", size = 12, face = "bold"), 
    axis.text.x = element_text(colour = "black", face = "bold", size = 20), 
    legend.text = element_text(size = 12, face ="bold", colour ="black"), 
    legend.position = "right", axis.title.y = element_text(face = "bold", size = 20), 
    axis.title.x = element_text(face = "bold", size = 12, colour = "black"), 
    legend.title = element_text(size = 12, colour = "black", face = "bold"), 
    panel.background = element_blank(), panel.border = element_rect(colour = "black", fill = NA, size = 1.2),
    legend.key=element_blank()) + 
    labs(x = "nmds_result", colour = "Treatment", y = "NMDS2", shape = "Site")  + 
    scale_colour_manual(values = c("#009E73", "#E69F00")) 

plot1


#plot2
co=c("red","black") 
shape=c(18,17,16) 
plot(nmds_result$points, col=interiorsurvey$Treatment, pch = interiorsurvey$Site, cex = 0.5, main = "Percent cover per treatment per site", xlab = "axis 1", ylab = "axis 2") 
  Site <- c("Jungle","Dunes") 
  legend('bottomright', Site, pch = c(), cex = 0.5, bty = "y" ) 
  Treatment <- c("Sargassum", "Control") 
  legend('bottomleft', Treatment, col=c("red","black"), pch = c(16), cex = 0.5, bty = "y" )



###diversity index




survey <- (interiorsurvey[,c(1,3,4,5,6,8:60)])
survey$Block<-as.factor(survey$Block)
survey$Quadrat<-as.factor(survey$Quadrat)
survey$Trip<-as.factor(survey$Trip)



###dune index



survey <- (interiorsurvey[,c(1,3,4,5,6,8:60)])
survey$Block<-as.factor(survey$Block)
survey$Quadrat<-as.factor(survey$Quadrat)
survey$Trip<-as.factor(survey$Trip)

Dunes_survey_data<-subset(survey, Site == "Dunes")
#merge litter and dead since litter was supposed to be categorized as dead

Dunes_survey_data$Dead_Lit <- paste(Dunes_survey_data$Dead + Dunes_survey_data$Litter)

# Remove the original columns if needed
Dunes_survey_data <- Dunes_survey_data[, !(names(Dunes_survey_data) %in% c("Dead", "Litter"))]


# Get the column names from the 6th column onwards
Dunes_species_columns <- names(Dunes_survey_data)[6:ncol(Dunes_survey_data)]

# Filter out columns with only zeros
Dunes_non_zero_columns <- Dunes_species_columns[!apply(Dunes_survey_data[Dunes_species_columns], 2, function(x) all(x == 0))]

# Create a new data frame with the non-zero columns
Dunes_filtered_df <- Dunes_survey_data[c('Trip', 'Site', 'Block', 'Treatment', 'Quadrat', Dunes_non_zero_columns)]
view(Dunes_filtered_df)


# Convert the columns to numeric if they are not already
Dunes_filtered_df[, Dunes_non_zero_columns] <- lapply(Dunes_filtered_df[, Dunes_non_zero_columns], as.numeric)

# Convert numbers to percent cover for each row independently
for (row in 1:nrow(Dunes_filtered_df)) {
  row_sum <- sum(Dunes_filtered_df[row, Dunes_non_zero_columns])
  if (row_sum != 0) {
    Dunes_filtered_df[row, Dunes_non_zero_columns] <- (Dunes_filtered_df[row, Dunes_non_zero_columns] / row_sum) * 100
  }
}

# Calculate Bray-Curtis dissimilarity matrix
Dunes_bray_curtis_dist <- vegdist(Dunes_filtered_df[, Dunes_non_zero_columns], method = "bray")

# Calculate beta diversity (average dissimilarity) using Bray-Curtis
Dunes_beta_diversity <- mean(Dunes_bray_curtis_dist)

# View the calculated beta diversity
Dunes_beta_diversity

# Assuming your dissimilarity matrix is named 'bray_curtis_dist' and you have the corresponding factors in 'Trip', and 'Treatment'

# Perform PERMANOVA to compare beta diversity
Dunes_permanova_result <- adonis2(Dunes_bray_curtis_dist ~ Treatment/Trip, data = Dunes_filtered_df)


# View the PERMANOVA results
Dunes_permanova_result


###jungles index


survey <- (interiorsurvey[,c(1,3,4,5,6,8:60)])
survey$Block<-as.factor(survey$Block)
survey$Quadrat<-as.factor(survey$Quadrat)
survey$Trip<-as.factor(survey$Trip)
Jungle_survey_data <- subset(survey, Site == "Jungle")

#merge litter and dead since litter was supposed ot be categorized as dead
Jungle_survey_data$Dead_Lit <- paste(Jungle_survey_data$Dead + Jungle_survey_data$Litter)
# Remove the original columns if needed
Jungle_survey_data <- Jungle_survey_data[, !(names(Jungle_survey_data) %in% c("Dead", "Litter"))]
# Get the column names from the 6th column onwards
Jungle_species_columns <- names(Jungle_survey_data)[6:ncol(Jungle_survey_data)]
view(Jungle_species_columns)
# Filter out columns with only zeros
Jungle_non_zero_columns <- Jungle_species_columns[!apply(Jungle_survey_data[Jungle_species_columns], 2, function(x) all(x == 0))]
# Create a new data frame with the non-zero columns
Jungle_filtered_df <- Jungle_survey_data[c('Trip', 'Site', 'Block', 'Treatment', 'Quadrat', Jungle_non_zero_columns)]
view(Jungle_filtered_df)
# Convert the columns to numeric if they are not already
Jungle_filtered_df[, Jungle_non_zero_columns] <- lapply(Jungle_filtered_df[, Jungle_non_zero_columns], as.numeric)
# Convert numbers to percent cover for each row independently
for (row in 1:nrow(Jungle_filtered_df)) {
  row_sum <- sum(Jungle_filtered_df[row, Jungle_non_zero_columns])
  if (row_sum != 0) {
    Jungle_filtered_df[row, Jungle_non_zero_columns] <- (Jungle_filtered_df[row, Jungle_non_zero_columns] / row_sum) * 100
  }
}
# Calculate Bray-Curtis dissimilarity matrix
Jungle_bray_curtis_dist <- vegdist(Jungle_filtered_df[, Jungle_non_zero_columns], method = "bray")
# Calculate beta diversity (average dissimilarity) using Bray-Curtis
Jungle_beta_diversity <- mean(Jungle_bray_curtis_dist)
# View the calculated beta diversity
Jungle_beta_diversity
# Assuming your dissimilarity matrix is named 'bray_curtis_dist' and you have the corresponding factors in 'Trip', and 'Treatment'
# Perform PERMANOVA to compare beta diversity
Jungle_permanova_result <- adonis2(Jungle_bray_curtis_dist ~ Treatment/Trip, data = Jungle_filtered_df)
# View the PERMANOVA results
Jungle_permanova_result
#all significant but we should focus on seeing if there is a treatment effect within each site rather than by between

```

##Ladder

```{r}
#import dataset
ladder<-read.csv("ladder.csv", header = TRUE)
ladder$Trip<-as.factor(ladder$Trip)
ladder$Quadrat<-as.factor(ladder$Quadrat)

#make data long 
laddersurvey_long <- gather(ladder, key = "cover_type", value = "percent_cover", 16:17, -c(1:5),) 
#use this for graphing purposes
laddersurvey_long$percent_cover<-as.numeric(laddersurvey_long$percent_cover)
laddersurvey_long$Trip<-as.factor(laddersurvey_long$Trip)

#graph it
#lets look at the distance effect of sargassum on percent cover of grass and all other plants in the jungle 


ladder<- laddersurvey_long %>%
  ggplot() +
  aes(x = Quadrat, y = percent_cover, fill = cover_type) +
  stat_summary(
    fun.y = "mean",
    geom = "bar",
    position = "dodge",
    width = 0.7
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.7),
    width = 0.25
  ) +
  scale_fill_hue(direction = 1) +
  labs(
    x = "Distance from plot (m)",
    y = "Percent cover",
    fill = "Cover type") +
  theme_minimal() +
  facet_grid(vars(Trip), vars(Site), scales = "free")

ladder

```



#Arthropods
##sticky trapexploration

```{r}

stickies <-read.csv("stickies.csv",header=T)
#remove total column

stickies <- subset(stickies, select = -Total)

#subset dunes 
stickies_dunes<- subset(stickies_data_long, Site == "Dune")

ggplot(stickies_dunes) +
 aes(x = Order, fill = Treatment, weight = Count) +
 geom_bar(position = "dodge") +
 scale_fill_hue(direction = 1) +
 theme_minimal() +
 facet_wrap(vars(Trip), scales = "free", nrow = 4L) +
    labs(title = "Aerial Arthropod Counts at Beach")

#subset jungle
stickies_jungle<- subset(stickies_data_long, Site == "Jungle")

ggplot(stickies_jungle) +
 aes(x = Order, fill = Treatment, weight = Count) +
 geom_bar(position = "dodge") +
 scale_fill_hue(direction = 1) +
 theme_minimal() +
 facet_wrap(vars(Trip), scales = "free", nrow = 4L) +
    labs(title = "Aerial Arthropod Counts at Jungle")

#explore Jungle Data
#total counts bar graph

stickies_jungle %>%
 filter(Order %in% c("Orthoptera", "Hemiptera", "Arachnid", "Diptera", "Hymenoptera", 
"Lepidoptera", "Empicoris", "Odonata")) %>%
 ggplot() +
  aes(x = Order, fill = Treatment, weight = Count) +
  geom_bar(position = "dodge") +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_wrap(vars(Trip), scales = "free", nrow = 4L) +
    labs(title = "Main contributors to flying arthropods counts in Jungle")

#the three main contributors to variability 
stickies_jungle %>%
 filter(Order %in% c("Hemiptera", "Diptera", "Hymenoptera")) %>%
 ggplot() +
  aes(x = Order, fill = Treatment, weight = Count) +
  geom_bar(position = "dodge") +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_wrap(vars(Trip), scales = "free", nrow = 4L) 

#now as a box plot 

stickies_jungle %>%
 filter(Order %in% c("Hemiptera", "Diptera", "Hymenoptera")) %>%
 ggplot() +
  aes(x = Order, y = Count, fill = Treatment) +
  geom_boxplot() +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_wrap(vars(Trip), scales = "free", nrow = 4L)
#explore Dune Data
#total counts bar graph with free scales after removing arthropods that are crawlers, don’t add much to the graphs, or pollinators.
stickies_dunes %>%
 filter(Order %in% c("Amphipods", "Coleoptera", "Hemiptera", "Diptera", "Hymenoptera"
)) %>%
 ggplot() +
  aes(x = Order, fill = Treatment, weight = Count) +
  geom_bar(position = "dodge") +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_wrap(vars(Trip), scales = "free", nrow = 4L)

# the pattern looks more pronounced and maybe reads easiest with fixed scales

stickies_dunes %>%
 filter(Order %in% c("Amphipods", "Coleoptera", "Hemiptera", "Diptera", "Hymenoptera"
)) %>%
 ggplot() +
  aes(x = Order, fill = Treatment, weight = Count) +
  geom_bar(position = "dodge") +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_wrap(vars(Trip), nrow = 4L) +
    labs(title = "Main contributors to flying arthropods counts on Dunes")
#the fly pattern is so interesting! No doubt of a subsidy effect there. 

#now as a boxplot
stickies_dunes %>%
 filter(Order %in% c("Amphipods", "Coleoptera", "Hemiptera", "Diptera", "Hymenoptera"
)) %>%
 ggplot() +
  aes(x = Order, y = Count, fill = Treatment) +
  geom_boxplot() +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_wrap(vars(Trip), nrow = 4L) +
    labs(title = "Main contributors to flying arthropods counts on Dunes")

#pods only… as expected, we see a surge on initial day and nothing much after

stickies_dunes %>%
 filter(Order %in% "Amphipods") %>%
 ggplot() +
  aes(x = Order, y = Count, fill = Treatment) +
  geom_boxplot() +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_wrap(vars(Trip), scales = "free", nrow = 4L)



```


### sticky trap analysis

```{r}
#load data from stickies1.csv, i curated that data sheet to only include the 3 orders we are interested in, added side 1 to side 2 of each trap, and simple averaged averaged "hi" and "lo" for each block/treatment. I decided to do this in excel in the interest of time.

stickiesmod1 <-read.csv("stickies1.csv",header=T)

############################################dunes############################################
#subset out the dunes
eff_sub_dune<-subset(stickiesmod1, Site == "Dune")

#subset trip 1
eff_sub_dune_1<-subset(eff_sub_dune, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
dune_df_sargassum <- eff_sub_dune_1[eff_sub_dune_1$Treatment == "Sargassum", ]
dune_df_control <- eff_sub_dune_1[eff_sub_dune_1$Treatment == "Control", ]

### effect sizes for trip 1
cohen.d(dune_df_sargassum$Amphipods,dune_df_control$Amphipods)
cohen.d(dune_df_sargassum$Diptera,dune_df_control$Diptera)
cohen.d(dune_df_sargassum$Hymenoptera,dune_df_control$Hymenoptera)

#subset trip 2
eff_sub_dune_2<-subset(eff_sub_dune, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
dune_df_sargassum2 <- eff_sub_dune_2[eff_sub_dune_2$Treatment == "Sargassum", ]
dune_df_control2 <- eff_sub_dune_2[eff_sub_dune_2$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(dune_df_sargassum2$Amphipods,dune_df_control2$Amphipods)
cohen.d(dune_df_sargassum2$Diptera,dune_df_control2$Diptera)
cohen.d(dune_df_sargassum2$Hymenoptera,dune_df_control2$Hymenoptera)

#subset trip 3
eff_sub_dune_3<-subset(eff_sub_dune, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
dune_df_sargassum3 <- eff_sub_dune_3[eff_sub_dune_3$Treatment == "Sargassum", ]
dune_df_control3 <- eff_sub_dune_3[eff_sub_dune_3$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(dune_df_sargassum3$Amphipods,dune_df_control3$Amphipods)
cohen.d(dune_df_sargassum3$Diptera,dune_df_control3$Diptera)
cohen.d(dune_df_sargassum3$Hymenoptera,dune_df_control3$Hymenoptera)


############################################jungle############################################
#subset out the dunes
eff_sub_jungle<-subset(stickiesmod1, Site == "Jungle")

#subset trip 1
eff_sub_jungle_1<-subset(eff_sub_jungle, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
jungle_df_sargassum <- eff_sub_jungle_1[eff_sub_jungle_1$Treatment == "Sargassum", ]
jungle_df_control <- eff_sub_jungle_1[eff_sub_jungle_1$Treatment == "Control", ]

### effect sizes for trip 1
cohen.d(jungle_df_sargassum$Amphipods,jungle_df_control$Amphipods)
cohen.d(jungle_df_sargassum$Diptera,jungle_df_control$Diptera)
cohen.d(jungle_df_sargassum$Hymenoptera,jungle_df_control$Hymenoptera)

#subset trip 2
eff_sub_jungle_2<-subset(eff_sub_jungle, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
jungle_df_sargassum2 <- eff_sub_jungle_2[eff_sub_jungle_2$Treatment == "Sargassum", ]
jungle_df_control2 <- eff_sub_jungle_2[eff_sub_jungle_2$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(jungle_df_sargassum2$Amphipods,jungle_df_control2$Amphipods)
cohen.d(jungle_df_sargassum2$Diptera,jungle_df_control2$Diptera)
cohen.d(jungle_df_sargassum2$Hymenoptera,jungle_df_control2$Hymenoptera)

#subset trip 3
eff_sub_jungle_3<-subset(eff_sub_jungle, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
jungle_df_sargassum3 <- eff_sub_jungle_3[eff_sub_jungle_3$Treatment == "Sargassum", ]
jungle_df_control3 <- eff_sub_jungle_3[eff_sub_jungle_3$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(jungle_df_sargassum3$Amphipods,jungle_df_control3$Amphipods)
cohen.d(jungle_df_sargassum3$Diptera,jungle_df_control3$Diptera)
cohen.d(jungle_df_sargassum3$Hymenoptera,jungle_df_control3$Hymenoptera)

#####Effect sizes Graph#####
#I manually input the effect sizes, lower and upper limits to a file called stickies_effsize.csv
stickieseffsize <-read.csv("sticky_effsize.csv",header=T)

stickieseffsize$Trip<-as.factor(stickieseffsize$Trip)



sticky_eff_plot <- ggplot(stickieseffsize, aes(x = Trip, y = Effect_size, shape = Order)) +
  labs(y = "Effect size", shape = "Taxa", title = "Aerial traps") +
  geom_path(aes(group = Order), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L, labeller = as_labeller(c("Dune" = "Beach", "Jungle" = "Forest"))) +
     geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
    scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "", "2" = "", "3" = "") ) +
  theme(
    axis.text = element_text(size=12),
    axis.title = element_text(size=12),
    plot.title = element_text(size=12),
    axis.text.x = element_text(size=12),
    strip.text = element_text(size=12)) +
   xlab("") 
#geom_pointrange(aes(ymin = lower_ci, ymax = upper_ci), position = position_dodge(0.5), color = " gray") +

sticky_eff_plot


```

###sticky trap figures

```{r}


#load data
stickies <-read.csv("stickies1.csv",header=T)  
stickies$Trip<-as.factor(stickies$Trip)

# Calculate means and standard errors for each combination of Trip, Taxa, and Treatment
stickies_data <- stickies %>%
  group_by(Trip, Site, Treatment, Taxa) %>%
  summarize(mean_count = mean(Count), 
            std_error = sd(Count) / sqrt(n()))

# Create the bar plot with means and error bars
ggplot(stickies_data) +
  aes(x = Trip, y = mean_count, fill = Taxa) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5) +
  geom_errorbar(aes(ymin = mean_count - std_error, ymax = mean_count + std_error), 
                position = position_dodge(width = 0.9), width = 0.25, color = "black") +
  scale_fill_manual(values = c("white", "gray60", "gray40")) +
  theme_bw() + 
   scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    strip.background = element_blank() )+
  xlab("") +
  facet_grid(vars(Treatment), vars(Site), scales = "free_y")


#####Effect sizes Graph#####
#I manually input the effect sizes, lower and upper limits to a file called pitfall_effsize.csv
stickieseffsize <-read.csv("sticky_effsize.csv",header=T)

stickieseffsize$Trip<-as.factor(stickieseffsize$Trip)

stickies_eff_plot <- ggplot(stickieseffsize, aes(x = Trip, y = Effect_size, shape = Order)) +
  labs(y = "Effect size", shape = "Taxa", title = "stickie traps") +
  geom_path(aes(group = Order), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L, labeller = as_labeller(c("Dune" = "Beach", "Jungle" = "Forest"))) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    strip.text = element_blank(),  # Remove facet title
    strip.background = element_blank()  # Remove facet boxes
    # geom_pointrange(aes(ymin = lower_ci, ymax = upper_ci), position = position_dodge(0.5), color = "gray") +
  ) +
  xlab("")

stickies_eff_plot


```

#pitfalls
####pitfalls analysis

```{r}

pitfallsmod1 <-read.csv("Pitfalls.csv",header=T)

############################################dunes############################################
#subset out the dunes
piteff_sub_dune<-subset(pitfallsmod1, Site == "Dune")

#subset trip 1
piteff_sub_dune_1<-subset(piteff_sub_dune, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
pitdune_df_sargassum <- piteff_sub_dune_1[piteff_sub_dune_1$Treatment == "Sargassum", ]
pitdune_df_control <- piteff_sub_dune_1[piteff_sub_dune_1$Treatment == "Control", ]

### effect sizes for trip 1
cohen.d(pitdune_df_sargassum$Amphipods,pitdune_df_control$Amphipods)
cohen.d(pitdune_df_sargassum$Arachnid,pitdune_df_control$Arachnid)
cohen.d(pitdune_df_sargassum$Hymenoptera,pitdune_df_control$Hymenoptera)

#subset trip 2
piteff_sub_dune_2<-subset(eff_sub_dune, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
pitdune_df_sargassum2 <- piteff_sub_dune_2[piteff_sub_dune_2$Treatment == "Sargassum", ]
pitdune_df_control2 <- piteff_sub_dune_2[piteff_sub_dune_2$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(pitdune_df_sargassum2$Amphipods,pitdune_df_control2$Amphipods)
cohen.d(pitdune_df_sargassum2$Arachnid,pitdune_df_control2$Arachnid)
cohen.d(pitdune_df_sargassum2$Hymenoptera,pitdune_df_control2$Hymenoptera)

#subset trip 3
piteff_sub_dune_3<-subset(piteff_sub_dune, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
pitdune_df_sargassum3 <- piteff_sub_dune_3[piteff_sub_dune_3$Treatment == "Sargassum", ]
pitdune_df_control3 <- piteff_sub_dune_3[piteff_sub_dune_3$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(pitdune_df_sargassum3$Amphipods,pitdune_df_control3$Amphipods)
cohen.d(pitdune_df_sargassum3$Arachnid,pitdune_df_control3$Arachnid)
cohen.d(pitdune_df_sargassum3$Hymenoptera,pitdune_df_control3$Hymenoptera)


############################################jungle############################################
#subset out the dunes
piteff_sub_jungle<-subset(pitfallsmod1, Site == "Jungle")

#subset trip 1
piteff_sub_jungle_1<-subset(piteff_sub_jungle, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
pitjungle_df_sargassum <- piteff_sub_jungle_1[piteff_sub_jungle_1$Treatment == "Sargassum", ]
pitjungle_df_control <- piteff_sub_jungle_1[piteff_sub_jungle_1$Treatment == "Control", ]

### effect sizes for trip 1
cohen.d(pitjungle_df_sargassum$Amphipods,pitjungle_df_control$Amphipods)
cohen.d(pitjungle_df_sargassum$Arachnid,pitjungle_df_control$Arachnid)
cohen.d(pitjungle_df_sargassum$Hymenoptera,pitjungle_df_control$Hymenoptera)

#subset trip 2
piteff_sub_jungle_2<-subset(piteff_sub_jungle, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
pitjungle_df_sargassum2 <- piteff_sub_jungle_2[piteff_sub_jungle_2$Treatment == "Sargassum", ]
pitjungle_df_control2 <- piteff_sub_jungle_2[piteff_sub_jungle_2$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(pitjungle_df_sargassum2$Amphipods,pitjungle_df_control2$Amphipods)
cohen.d(pitjungle_df_sargassum2$Arachnid,pitjungle_df_control2$Arachnid)
cohen.d(pitjungle_df_sargassum2$Hymenoptera,pitjungle_df_control2$Hymenoptera)

#subset trip 3
piteff_sub_jungle_3<-subset(piteff_sub_jungle, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
pitjungle_df_sargassum3 <- piteff_sub_jungle_3[piteff_sub_jungle_3$Treatment == "Sargassum", ]
pitjungle_df_control3 <- piteff_sub_jungle_3[piteff_sub_jungle_3$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(pitjungle_df_sargassum3$Amphipods,pitjungle_df_control3$Amphipods)
cohen.d(pitjungle_df_sargassum3$Arachnid,pitjungle_df_control3$Arachnid)
cohen.d(pitjungle_df_sargassum3$Hymenoptera,pitjungle_df_control3$Hymenoptera)



#####Effect sizes#####
#I manually input the effect sizes, lower and upper limits to a file called pitfalls_effsize.csv
pitfallseffsize <-read.csv("pitfalls_effsize.csv",header=T)


sticky_eff_plot <- ggplot(pitfallseffsize) +
  aes(x = Trip, y = Effect_size, shape = Order) +
  geom_jitter(size = 1.2) +
  labs(y = "Effect size ", shape = "Taxa") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L)

sticky_eff_plot

```


###pitfall trap figures

```{r}
#load data
pitfalls <-read.csv("Pitfalls1.csv",header=T)  
pitfalls$Trip<-as.factor(pitfalls$Trip)

# Calculate means and standard errors for each combination of Trip, Taxa, and Treatment
pitfalls_data <- pitfalls %>%
  group_by(Trip, Site, Treatment, Taxa) %>%
  summarize(mean_count = mean(Count), 
            std_error = sd(Count) / sqrt(n()))

# Create the bar plot with means and error bars
ggplot(pitfalls_data) +
  aes(x = Trip, y = mean_count, fill = Taxa) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5) +
  geom_errorbar(aes(ymin = mean_count - std_error, ymax = mean_count + std_error), 
                position = position_dodge(width = 0.9), width = 0.25, color = "black") +
  scale_fill_manual(values = c("white", "gray60", "gray40")) +
  theme_bw() + 
   scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    strip.background = element_blank() )+
  xlab("") +
  facet_grid(vars(Treatment), vars(Site), scales = "free_y")


#####Effect sizes Graph#####
#I manually input the effect sizes, lower and upper limits to a file called pitfall_effsize.csv
pitfalleffsize <-read.csv("pitfalls_effsize.csv",header=T)

pitfalleffsize$Trip<-as.factor(pitfalleffsize$Trip)

pitfall_eff_plot <- ggplot(pitfalleffsize, aes(x = Trip, y = Effect_size, shape = Order)) +
  labs(y = "Effect size", shape = "Taxa", title = "Pitfall traps") +
  geom_path(aes(group = Order), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L, labeller = as_labeller(c("Dune" = "Beach", "Jungle" = "Forest"))) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    strip.text = element_blank(),  # Remove facet title
    strip.background = element_blank()  # Remove facet boxes
    # geom_pointrange(aes(ymin = lower_ci, ymax = upper_ci), position = position_dodge(0.5), color = "gray") +
  ) +
  xlab("")

pitfall_eff_plot



```

#Chemicals and gases
##nutrients

```{r}

##quick nutrients

chemicals<-read.csv("Soil_Chemicals2.csv",header=T)

#subset experiment
pile_chem<-subset(chemicals, Exp == "Sargasso Manipulation")
pile_chem$Block<-as.factor(pile_chem$Block)
pile_chem$Trip<-as.factor(pile_chem$Trip)
pile_chem$pH<-as.numeric(pile_chem$pH)
pile_chem$KH<-as.numeric(pile_chem$KH)


#explore
#nitrate by block at beach and forest for 4 trips
ggplot(pile_chem) +
  aes(x = Site, y = Nitrate, fill = Block) +
  geom_boxplot() +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_grid(vars(Trip), vars())




#added treatment. It is way too convoluted and impossible t read this way
pile_chem %>%
 filter(Type %in% "Sargassum") %>%
 ggplot() +
 aes(x = Site, y = Nitrate, fill = Trt, colour = Block) +
 geom_boxplot() +
 scale_fill_hue(direction = 1) +
 scale_color_hue(direction = 1) +
 theme_minimal() +
 facet_grid(vars(Trip), vars())


# We are interested in the sediment nutrients rather than sediment and seaweed


pile_chem %>%
  filter(Type %in% "Sargassum") %>%
  filter(Trt %in% "Sediment") %>%
  group_by(Trip, Site) %>%
  summarize(mean_Nitrate = mean(Nitrate), se = sd(Nitrate) / sqrt(n())) %>%
  ggplot() +
  aes(x = Trip, y = mean_Nitrate, fill = Site) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean_Nitrate - se, ymax = mean_Nitrate + se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  scale_fill_hue(direction = 1) +
    scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23") ) +
  labs(
    x = " ",
    y = "Nitrate content (mg/L)",
    title = "",
    fill = "Sample"
  ) +
  theme_minimal()  +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"))


```

##gasses
```{r}
CO2<-read.csv("gas_readings.csv",header=T)

#subset experiment
gas_CO2<-subset(CO2, !Treatment == "na")
gas_CO2$Carbon_dioxide<-as.numeric(gas_CO2$Carbon_dioxide)
gas_CO2$Trip<-is.factor(gas_CO2$Carbon_dioxide)



#gas collectors
gas_CO2 %>%
  filter(Experiment %in% "collectors") %>%
  ggplot() +
  aes(x = Treatment, y = Carbon_dioxide, fill = Time) +
  stat_summary(
    fun = "mean",
    geom = "bar",
    position = "dodge",
    width = 0.7
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.7),
    width = 0.25
  ) +
  scale_fill_hue(direction = 1) +
  labs(
    x = " ",
    y = "Units",
    title = "Gas collector",
    fill = "Time point"
  ) +
    theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    axis.title = element_text(size = 16),   # Adjust the size of axis titles
    axis.text = element_text(size = 14),    # Adjust the size of axis labels
    axis.text.x = element_text(size = 14),  # Adjust the size of x-axis labels
    plot.title = element_text(size = 18),   # Adjust the size of the plot title (if needed)
    legend.title = element_text(size = 16), # Adjust the size of the legend title
    legend.text = element_text(size = 14)   # Adjust the size of legend labels
  ) +
  facet_grid(vars(), vars(Site), scales = "free_y")

#respirometer

gas_CO2 %>%
  filter(Experiment %in% "respirometer") %>%
  ggplot() +
  aes(x = Treatment, y = Carbon_dioxide, fill = Position) +
  stat_summary(
    fun = "mean",
    geom = "bar",
    position = "dodge",
    width = 0.7
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.7),
    width = 0.25
  ) +
  scale_fill_hue(direction = 1) +
  labs(
    x = " ",
    y = "Units",
    title = "Respirometer",
    fill = "Reader position"
  ) +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    axis.title = element_text(size = 16),   # Adjust the size of axis titles
    axis.text = element_text(size = 14),    # Adjust the size of axis labels
    axis.text.x = element_text(size = 14),  # Adjust the size of x-axis labels
    plot.title = element_text(size = 18),   # Adjust the size of the plot title (if needed)
    legend.title = element_text(size = 16), # Adjust the size of the legend title
    legend.text = element_text(size = 14)   # Adjust the size of legend labels
  ) +
  facet_grid(vars(), vars(Site))



```


#Manuscript figures
##pile vol
```{r}

#####################################pile volumes#####################################

piles <-read.csv("Sargassum_piles.csv",header=T)
piles2<-subset(piles, Treatment== "Sargassum")
piles2$Trip<-as.factor(piles2$Trip)
piles2$Block<-as.factor(piles2$Block)
# Calculate the means by month and location

mean_data <- piles2 %>%
  group_by(Trip, Location) %>%
  summarize(mean_ellipsoid_vol = mean(Ellipsoid_prc_D))


# Create the scatter plot with mean points connected by a line
scatterplot <- ggplot(data = piles2, aes(x = Trip, y = Ellipsoid_prc_D, fill = Location)) +
  geom_point(size = 2, shape = 21, aes(fill = Location)) +
  geom_line(data = mean_data, aes(x = Trip, y = mean_ellipsoid_vol, group = Location), color = "gray")

# Customize the plot appearance
vol_x_time <- scatterplot +
  labs(x = "", y = "Pile volume (%)") +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_manual(values = c("black", "white")) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar.23", "4" = "Aug.23")
  ) +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    axis.title = element_text(size = 16),   # Adjust the size of axis titles
    axis.text = element_text(size = 14),    # Adjust the size of axis labels
    axis.text.x = element_text(size = 14),  # Adjust the size of x-axis labels
    plot.title = element_text(size = 18),   # Adjust the size of the plot title (if needed)
    legend.title = element_text(size = 16), # Adjust the size of the legend title
    legend.text = element_text(size = 14)   # Adjust the size of legend labels
  )

vol_x_time
```

##decomp 
```{r}

##############################decomposer experiment##################################
#now lets look at mean values and SE
#subset out trip 1 since its only going to show initial mass which is 100%
decomp %>%
  filter(!(Trip %in% "1")) %>%
  filter(!(ratio %in% "")) %>%
  ggplot() +
  scale_fill_manual(name = "Mesh size", 
                    values = c("large" = "light gray", "small" = "white")) +  # Set bars to white
  scale_color_manual(name = "Mesh size",
                     values = c("large" = "black", "small" = "black")) +  # Set outline color to black
  aes(x = Site, fill = Treatment, weight = prc_means) +  # Add color aesthetic
  geom_bar(position = "dodge", color = "black") +  # Specify outline color
  geom_errorbar(aes(ymin = prc_means - prc_SE, ymax = prc_means + prc_SE), 
                position = position_dodge(width = 0.9), 
                width = 0.25) +  
  labs(y = "Decomposition (% of initial dry biomass)", x = "") +
  theme_minimal() +
  facet_grid(Trip ~ ., labeller = labeller(Trip = function(x) ifelse(x == "2", "Nov. 22", ifelse(x == "3", "Mar. 23", x)))) + 
    theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    axis.title = element_text(size = 16),   # Adjust the size of axis titles
    axis.text = element_text(size = 14),    # Adjust the size of axis labels
    axis.text.x = element_text(size = 14),  # Adjust the size of x-axis labels
    plot.title = element_text(size = 18),   # Adjust the size of the plot title (if needed)
    legend.title = element_text(size = 16), # Adjust the size of the legend title
    legend.text = element_text(size = 14)   # Adjust the size of legend labels
  )
```
## nutrients 

```{r}
#############################nutrients##############################################
pile_chem %>%
  filter(Type %in% "Sargassum") %>%
  filter(Trt %in% "Sediment") %>%
  group_by(Trip, Site) %>%
  summarize(mean_Nitrate = mean(Nitrate), se = sd(Nitrate) / sqrt(n())) %>%
  ggplot() +
  aes(x = Trip, y = mean_Nitrate, fill = Site) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  geom_errorbar(aes(ymin = mean_Nitrate - se, ymax = mean_Nitrate + se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  scale_fill_manual(values = c("lightgray", "white")) +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
  labs(
    x = " ",
    y = "Nitrate content (mg/L)",
    title = "",
    fill = "Sample"
  ) +
  theme_minimal() +
 theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_blank()) # Remove facet labels

```

##CO2
```{r}
##############################gases#################################################
#gas collectors

CO2<-read.csv("gas_readings.csv",header=T)

#subset experiment
collectors<-subset(CO2, Experiment == "collectors") 
collectors<-subset(collectors, !Site == "Calibrate")


CO2plot1 <- collectors %>%
  group_by(Site, Treatment, Time) %>%
  summarize(mean_CO2 = mean(Carbon_dioxide), se = sd(Carbon_dioxide) / sqrt(n())) %>%
  ggplot() +
  aes(x = Treatment, y = mean_CO2, fill = Time) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  geom_errorbar(aes(ymin = mean_CO2 - se, ymax = mean_CO2 + se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  scale_fill_manual(values = c("lightgray", "white")) +
 
  labs(
    x = " ",
    y = "CO2 UNITS",
    title = "",
    fill = "Sample"
  ) +
  scale_x_discrete(labels = NULL) +  # Remove x-axis labels
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 18),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  ) +
   facet_grid(vars(), vars(Site), scales = "free_y")



CO2plot1





#respirometer

CO2<-read.csv("gas_readings.csv",header=T)


#subset experiment
respirometer<-subset(CO2, Experiment == "respirometer") 

CO2plot2 <- respirometer %>%
  group_by(Site, Treatment, Position) %>%
  summarize(mean_CO2 = mean(Carbon_dioxide), se = sd(Carbon_dioxide) / sqrt(n())) %>%
  ggplot() +
  aes(x = Treatment, y = mean_CO2, fill = Position) +
  geom_bar(stat = "identity", position = "dodge", color = "black") +
  geom_errorbar(aes(ymin = mean_CO2 - se, ymax = mean_CO2 + se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  scale_fill_manual(values = c("lightgray", "white")) +
  labs(
    x = " ",
    y = "CO2 UNITS",
    title = "",
    fill = "Position"
  ) +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  ) +
  facet_grid(vars(), vars(Site), scales = "free_y")


CO2plot2 <- CO2plot2 +
  theme(
    axis.title.y = element_text(margin = margin(r = 25)),  # Adjust the margin
    # Other theme settings...
  )

CO2plot2


combined_CO2 <- grid.arrange(CO2plot1, CO2plot2, ncol = 1)

```

##interior percent cover
```{r}
##############################interior percent cover##################################
interior_pc_cover<- ggplot(data = interiorsurvey2,
            aes(
              x = Trip,
              y = means,
              shape = cover_type,
              color = Treatment
            )) +
  stat_summary(geom = "point", fun = mean, position = position_dodge(width = 1),
               size = 2L, stroke = 1) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 1),
               width = 0.2) +  # Adjust the width of the error bars as needed
  labs(
    x = " ",
    y = "Mean percent cover",
    title = "Interior Mean Percent Cover and Effect Sizes",
    shape = "Cover type"
  ) +
   scale_shape_discrete(
    labels = c("grass", "other plants"),
    name = "Cover"  # Change the legend title to "Cover"
  ) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 23")
  ) +
  facet_grid(vars(), vars(Site)) +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_blank()  # Remove facet labels
  ) +
  scale_fill_manual(values = c("grey44", "black")) +
  scale_color_manual(values = c("grey44", "black")) 
  
  

interior_pc_cover

#interior effect size graph
int_eff_plot <- ggplot(interioreffsize, aes(x = Trip, y = Effect_size, shape = Cover)) +
  labs(y = "Effect size", title = "") +
  geom_path(aes(group = Cover), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar.23", "4" = "Aug.23")
  ) +
    scale_shape_discrete(labels = c("grass", "other plants")) +
 theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_blank()  # Remove facet labels
  ) +
  xlab("") 

int_eff_plot

#combine 
combined_int_effgraphs <- grid.arrange(interior_pc_cover, int_eff_plot, ncol = 1)

```

##perimeter percent cover
```{r}
#################################perimeter percent cover#########################
perimeter_pc_cover<- ggplot(data = perimetersurvey2,
            aes(
              x = Trip,
              y = means,
              shape = cover_type,
              color = Treatment
            )) +
  stat_summary(geom = "point", fun = mean, position = position_dodge(width = 1),
               size = 2L, stroke = 1) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 1),
               width = 0.2) +  # Adjust the width of the error bars as needed
  labs(
    x = " ",
    y = "Mean percent cover",
    title = "perimeter Mean Percent Cover and Effect Sizes",
    shape = "Cover type"
  ) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 23")
  ) +
    scale_shape_discrete(
    labels = c("grass", "other plants"),
    name = "Cover"  # Change the legend title to "Cover"
  ) +
theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  ) +
  facet_grid(vars(), vars(Site), scales = "free_y") +
  scale_fill_manual(values = c("grey44", "black")) +
  scale_color_manual(values = c("grey44", "black")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines")
  )

perimeter_pc_cover

#eff sizes 
per_eff_plot <- ggplot(perimetereffsize, aes(x = Trip, y = Effect_size, shape = Cover)) +
  labs(y = "Effect size", title = "") +
  geom_path(aes(group = Cover), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar.23", "4" = "Aug.23")
  ) +
  scale_shape_discrete(
    labels = c("grass", "other plants"),
    name = "Cover"  # Change the legend title to "Cover"
  ) +
theme_minimal() +
theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_blank()  # Remove facet labels
  ) +
  facet_grid(vars(), vars(Site), scales = "free_y") +
xlab("") 

per_eff_plot

#combine 
combined_perimeter_graph<- grid.arrange(perimeter_pc_cover, per_eff_plot, ncol = 1)

```


##arthropods
```{r}


##################################arthropods##########################

#pitfalls
#load data
pitfalls <-read.csv("Pitfalls1.csv",header=T)  
pitfalls$Trip<-as.factor(pitfalls$Trip)

# Calculate means and standard errors for each combination of Trip, Taxa, and Treatment
pitfalls_data <- pitfalls %>%
  group_by(Trip, Site, Treatment, Taxa) %>%
  summarize(mean_count = mean(Count), 
            std_error = sd(Count) / sqrt(n()))

# Create the bar plot with means and error bars
pitfall_prc_cover<-ggplot(pitfalls_data) +
  aes(x = Trip, y = mean_count, fill = Taxa) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5) +
  geom_errorbar(aes(ymin = mean_count - std_error, ymax = mean_count + std_error), 
                position = position_dodge(width = 0.9), width = 0.25, color = "black") +
  scale_fill_manual(values = c("white", "gray60", "gray40")) +
  theme_bw() + 
   scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_blank()  # Remove facet labels
  ) +
  xlab("") +
  ylab("Crawling arthropod count") +
  facet_grid(vars(Treatment), vars(Site), scales = "free_y")

pitfall_prc_cover

#####Effect sizes Graph#####
#I manually input the effect sizes, lower and upper limits to a file called pitfall_effsize.csv
pitfalleffsize <-read.csv("pitfalls_effsize.csv",header=T)

pitfalleffsize$Trip<-as.factor(pitfalleffsize$Trip)

pitfall_eff_plot <- ggplot(pitfalleffsize, aes(x = Trip, y = Effect_size, shape = Order)) +
  labs(y = "Effect size", shape = "Taxa", title = "Pitfall traps") +
  geom_path(aes(group = Order), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L, labeller = as_labeller(c("Dune" = "Beach", "Jungle" = "Forest"))) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_blank()  # Remove facet labels
  ) +
  xlab("")

pitfall_eff_plot

combined_pitfall_graphs <- grid.arrange(pitfall_prc_cover, pitfall_eff_plot, ncol = 1)




#aerial traps


#load data
stickies <-read.csv("stickies1.csv",header=T)  
stickies$Trip<-as.factor(stickies$Trip)

# Calculate means and standard errors for each combination of Trip, Taxa, and Treatment
stickies_data <- stickies %>%
  group_by(Trip, Site, Treatment, Taxa) %>%
  summarize(mean_count = mean(Count), 
            std_error = sd(Count) / sqrt(n()))

# Create the bar plot with means and error bars
stickies_prc_cover<-ggplot(stickies_data) +
  aes(x = Trip, y = mean_count, fill = Taxa) +
  geom_bar(stat = "identity", position = "dodge", color = "black", size = 0.5) +
  geom_errorbar(aes(ymin = mean_count - std_error, ymax = mean_count + std_error), 
                position = position_dodge(width = 0.9), width = 0.25, color = "black") +
  scale_fill_manual(values = c("white", "gray60", "gray40")) +
  theme_bw() + 
   scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_blank()  # Remove facet labels
  ) +
  xlab("") +
    ylab("Flying arthropod count") +
  facet_grid(vars(Treatment), vars(Site), scales = "free_y")
stickies_prc_cover

#####Effect sizes Graph#####
#I manually input the effect sizes, lower and upper limits to a file called pitfall_effsize.csv
stickieseffsize <-read.csv("sticky_effsize.csv",header=T)

stickieseffsize$Trip<-as.factor(stickieseffsize$Trip)

stickies_eff_plot <- ggplot(stickieseffsize, aes(x = Trip, y = Effect_size, shape = Order)) +
  labs(y = "Effect size", shape = "Taxa", title = "stickie traps") +
  geom_path(aes(group = Order), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L, labeller = as_labeller(c("Dune" = "Beach", "Jungle" = "Forest"))) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_blank()  # Remove facet labels
  ) +
  xlab("")

stickies_eff_plot


combined_stickies_graphs <- grid.arrange(stickies_prc_cover, stickies_eff_plot, ncol = 1)

all_arthropod_graphs <- grid.arrange(combined_pitfall_graphs,combined_stickies_graphs, ncol = 2)

```
#Supplementary graphs
## Full ladder survey

```{r}

laddersurvey_long_sarg<-subset(laddersurvey_long, Treatment == "Sargassum")

ggplot(laddersurvey_long_sarg) +
  aes(x = Trip, fill = cover_type, y = percent_cover) +
    stat_summary(
    fun.y = "mean",
    geom = "bar",
    position = "dodge",
    width = 0.7   ) +
     scale_fill_manual(values = c("lightgray", "darkgray"),  # Specify custom colors
                    labels = c("grass", "other plants")) +     # Specify custom labels
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.7),
    width = 0.25
  ) +
  labs(
    x = "Distance from plot (m)",
    y = "Percent cover",
    fill = "Cover type") +
    scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
   labs(
    x = "",
    y = "Percent cover",
    fill = "Cover type") +
  theme_minimal() +
   theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 24),
    axis.title = element_text(size = 26),
    axis.text = element_text(size = 24),
    axis.text.x = element_text(size = 24),
    plot.title = element_text(size = 22),
    legend.title = element_text(size = 26),
    legend.text = element_text(size = 24)
  ) +

  
  facet_grid(vars(Quadrat), vars(Site))

```
## Full perimeter percent cover


```{r}
#import dataset
perimetersurvey_full<-read.csv("perimeter_survey.csv", header = TRUE)

perimeter_prc_long<-perimetersurvey_full %>%
  gather(key = "cover_type", value = "prc_cover", c(7:58))

perimeter_prc_long$Trip<-as.factor(perimeter_prc_long$Trip) 
#remove zeros so the data is not so cluttered
# Remove rows where B is zero
perimeter_prc_long <- subset(perimeter_prc_long, prc_cover != 0)

view(perimeter_prc_long)
esquisse::esquisser(perimeter_prc_long)


# Create a heatmap
ggplot(perimeter_prc_long, aes(x = Trip, y = cover_type, fill = prc_cover)) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "blue") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))





perimetersurvey2<-subset(perimetersurvey, Data == "graph")
perimeter_total_pc_cover<- ggplot(data = perimetersurvey2,
            aes(
              x = Trip,
              y = means,
              shape = cover_type,
              color = Treatment
            )) +
  stat_summary(geom = "point", fun = mean, position = position_dodge(width = 1),
               size = 2L, stroke = 1) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 1),
               width = 0.2) +  # Adjust the width of the error bars as needed
  labs(
    x = " ",
    y = "Mean percent cover",
    title = "perimeter Mean Percent Cover and Effect Sizes",
    shape = "Cover type"
  ) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 23")
  ) +
    scale_shape_discrete(
    labels = c("grass", "other plants"),
    name = "Cover"  # Change the legend title to "Cover"
  ) +
theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14)
  ) +
  facet_grid(vars(), vars(Site), scales = "free_y") +
  scale_fill_manual(values = c("grey44", "black")) +
  scale_color_manual(values = c("grey44", "black")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines")
  )



```

```{r}
sargman<-read.csv("nmec.csv",header=T)

# Assuming your_data is your original data frame
sargman2 <- sargman %>%
  pivot_longer(cols = c(top_layer, bottom_layer), names_to = "layer", values_to = "species")










# Assuming your_data is your data frame
wide_data <- reshaped_data %>%
  pivot_wider(names_from = species,
              values_from = species,
              values_fn = length,
              values_fill = 0,
              names_prefix = "")

# Print the resulting wide_data
view(wide_data)


long_data <- gather(wide_data, key = "species", value = "count", 14:81, -c(1:13))


# Group the data by quadrat_code and survey_date, then calculate the sum of other columns
collapsed_df <- long_data %>%
  group_by(quadrat_code, survey_date) %>%
  summarize(total = sum(count),   # Replace 'column1' with the column you want to sum
            # Add more columns you want to sum here
            )

# Replace 'column1', etc. with the actual column names you want to sum.


# Group the data by 'quadrat_code' and 'survey_date', and summarize the data
summary_df <- wide_data %>%
  group_by(quadrat_code, survey_date) %>%
  summarize(total = sum(column_to_sum),  # Replace 'column_to_sum' with the column you want to summarize
            mean_value = mean(another_column))  # Replace 'another_column' with another column you want to summarize

# Replace 'column_to_sum' and 'another_column' with the actual columns you want to summarize.



# Assuming your data frame is named 'df'
library(dplyr)

# Add a new column 'total' that contains the sum of columns 3 through 30
wide_data2 <- wide_data %>%
  mutate(total = rowSums(across(14:81)))

# This calculates the sum of columns 3 through 30 for each row.

view(wide_data2)

```

