---
title: "Sargasso_Subisidy_Code"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options:
  chunk_output_type: console
---

#libraries

```{r}
library(knitr)
library(patchwork)
library(tidyverse)
library(tidyr)
library(ggplot2)
library(ggfortify)
library(lme4)
library(lmerTest)
library(car)
library(nlme)
library(ggpubr)
library(multcompView)
library(boot)
library(Hotelling)
library(mvnTest)
library(vegan)
library(factoextra)
library(FactoMineR)
library(agricolae)
library(glmm)
library(HH)
library(ggpubr)
library(rstatix)
library(vegan)
library(effsize)
library(broom)
library(performance)
library(DHARMa)
library(glmmTMB)
library(emmeans)
library(kableExtra)

setwd("~/Documents/GitHub/Sargasso_Subsidy")
#hello

```


# Pile metrics

```{r}
piles <-read.csv("Sargassum_piles.csv",header=T)
piles2<-subset(piles, Treatment== "Sargassum")
piles2$Trip<-as.factor(piles2$Trip)
piles2$Block<-as.factor(piles2$Block)

# Create the box plot
ggplot(piles2, aes(x = Trip, y = ellipsoid_vol, color = Location)) +
  geom_boxplot() 

#histogram
hist(piles2$ellipsoid_vol, main = "Histogram of Variable", xlab = "Variable Values")

# Create a Q-Q plot to assess normality
qqnorm(piles2$ellipsoid_vol)
qqline(piles2$ellipsoid_vol)


# Perform Levene's test for homogeneity of variances
levene_test <- leveneTest(ellipsoid_vol ~ Location, data = piles2)

print(levene_test)


#test if starting size of piles were different initially
#subset trip 1

view(piles3)
piles3<-subset(piles2, Trip == "1")

t.test(ellipsoid_vol~Location, piles3)

#since piles differed in size between sites and we are interested in their decomposition at each site, we will analyze our data as a percent of their respective initial volumes rather than total volume comparison. Thus initial starting volumes will be 100% across sites and need not be discussed further.

#Pile volume as the response, site as the main between-subjects factor, sampling date as the repeated within-subject 

piles_model <- lmer(Ellipsoid_prc_D ~ Trip * Location  + (1|Block), data = piles2)

summary(piles_model)
anova(piles_model)

#assumptions
plot(piles_model)

hist(resid(piles_model))
qqnorm(resid(piles_model))
qqline(resid(piles_model))
#data look normal enough

# Check for autocorrelation
acf(resid(piles_model))

# Extract residuals n fitted values
residuals <- resid(piles_model)
fitted_values <- fitted(piles_model)

# residuals vs fitted values
plot(fitted_values, residuals, 
     xlab = "Fitted values", ylab = "Residuals",
     main = "Residuals vs Fitted Values")

# Add line
abline(h = 0, col = "red")

#smooth
lines(stats::lowess(fitted_values, residuals), col = "blue")

#no obvious patterns, variance is fine. we accept the model.


```

### pile figures
```{r}
piles <-read.csv("Sargassum_piles.csv",header=T)
piles2<-subset(piles, Treatment== "Sargassum")
piles2$Trip<-as.factor(piles2$Trip)
piles2$Block<-as.factor(piles2$Block)

# Calculate the means by month and location

mean_data <- piles2 %>%
  group_by(Trip, Location) %>%
  summarize(mean_ellipsoid_vol = mean(Ellipsoid_prc_D))


# Create the scatter plot with mean points connected by a line
vol_scatterplot <- ggplot(data = piles2, aes(x = Trip, y = Ellipsoid_prc_D, fill = Location)) +
  geom_point(size = 2, shape = 21, aes(fill = Location)) +
  geom_line(data = mean_data, aes(x = Trip, y = mean_ellipsoid_vol, group = Location), color = "gray")

# Customize the plot a
vol_x_time<- vol_scatterplot + 
  labs(x = "", y = "Pile volume (%)") +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_manual(values = c("black", "white")) +
   scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar.23", "4" = "Aug.23")
  ) +
  theme_minimal() + 
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"))
vol_x_time



```


#Decomposer experiment 
##inverts

```{r}

decomp<-read.csv("mesh_bag_inverts.csv",header = T )
decomp$Rep_no.<-as.factor(decomp$Rep_no.)
decomp$Trip<-as.factor(decomp$Trip)

view(decomp)

decomp_inverts<- decomp %>%
  gather(key = "inverts", value = "count", c(5,6,11))

#remove small bags 
decomp_inverts2<-subset(decomp_inverts, Treatment == "large")


#figure out means and SE for graph

decomp_inverts2 %>%
  group_by(Trip, Site, Treatment, inverts) %>%
  summarize(mean_count = mean(count), count_se = sd(count) / sqrt(n())) %>%
  ggplot() +
  aes(x =inverts, y = mean_count, fill = Treatment) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean_count - count_se, ymax = mean_count + count_se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  scale_fill_hue(direction = 1) +
 theme_classic() +
  facet_grid(vars(Trip), vars(Site)) +
   theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14, color = "black"),
    axis.title = element_text(size = 16, color = "black"),
    axis.text = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    plot.title = element_text(size = 12, color = "black"),
    legend.title = element_text(size = 16, color = "black"),
    legend.text = element_text(size = 14, color = "black"),
    strip.text = element_text(size = 14, color = "black")) +  
  labs(y = "Arthroopod counts", x = "", title = "Mesh bag inverts")


# Calculate the total counts for each group to see how much higher amphipod counts are compared to others at beach on aug 2022
#subset aug 2022
decomp_inverts_aug<-subset(decomp, Trip == "Aug. 22")
#beach
decomp_inverts_aug_beach<- subset(decomp_inverts_aug, Site == "Beach")

total_amphipods <- sum(decomp_inverts_aug_beach$Amphipoda, na.rm = TRUE)
total_arachnids <- sum(decomp_inverts_aug_beach$Arachnida, na.rm = TRUE)
total_hymenopterans <- sum(decomp_inverts_aug_beach$Hymenoptera, na.rm = TRUE)

# Calculate the percentage difference for Amphipods compared to others
percent_difference <- ((total_amphipods - (total_arachnids + total_hymenopterans )) / (total_arachnids + total_hymenopterans )) * 100

# Print the results
cat("percent_difference:", percent_difference, "%\n")

#now the Forest
ecomp_inverts_aug_F<- subset(decomp_inverts_aug, Site == "Forest")

Ftotal_amphipods <- sum(ecomp_inverts_aug_F$Amphipoda, na.rm = TRUE)
Ftotal_arachnids <- sum(ecomp_inverts_aug_F$Arachnida, na.rm = TRUE)
Ftotal_hymenopterans <- sum(ecomp_inverts_aug_F$Hymenoptera, na.rm = TRUE)



# Fit linear mixed-effects model
#make a pair variable for site and replicate number for the random factor

decomp_inverts2$pair<-paste(sep = "",decomp_inverts2$Site,decomp_inverts2$Rep_no.)
view(decomp_inverts2)

decomp_inverts_model<- lm(count ~  inverts * Site * Trip , data = decomp_inverts2)


# Print model summary
summary(decomp_inverts_model)
anova(decomp_inverts_model)

# Obtain estimated marginal means
means <- emmeans(decomp_inverts_model, ~ inverts + Site * Trip)


# Perform pairwise comparisons
pairwise_comparisons <- pairs(means, by = "Site:Trip", infer = c(TRUE, TRUE))


# Print the comparisons
print(pairwise_comparisons)


```

##mesh bags
```{r}

#load data
decomp_bags<-read.csv("mesh_bag_exp.csv",header =T )
decomp_bags$Rep_no.<-as.factor(decomp_bags$Rep_no.)
decomp_bags$Trip<-as.factor(decomp_bags$Trip)
decomp_bags$prc_Delta <- as.numeric(decomp_bags$prc_Delta)


view(decomp_bags)


#quick view counts of percent
ggplot(decomp_bags) +
 aes(x = Trip, fill = Treatment, weight = prc_means) +
 geom_bar(position = "dodge") +
 scale_fill_hue(direction = 1) +
 theme_minimal() +
 facet_grid(vars(Site), vars())


#now lets look at mean values and SE
#subset out trip 1 since its only going to show initial mass which is 100%
decomp_bags %>%
  filter(!(Trip %in% "1")) %>%
  filter(!(ratio %in% "")) %>%
  ggplot() +
  scale_fill_manual(name = "Mesh size", 
                    values = c("large" = "grey25", "small" = "dark gray")) +
  aes(x = Site, fill = Treatment, weight = prc_means) +
  geom_bar(position = "dodge") +
  geom_errorbar(aes(ymin = prc_means - prc_SE, ymax = prc_means + prc_SE), 
                position = position_dodge(width = 0.9), 
                width = 0.25) +  
  labs(y = "Decomposition (% of initial dry biomass)") +
  theme_minimal() +
  facet_grid(Trip ~ ., labeller = labeller(Trip = function(x) ifelse(x == "2", "Nov. 22", ifelse(x == "3", "Mar. 23", x)))) 

view(decomp_bags)



#in our linear model, we want to see how percent decomposition differed by treatment at each site.  Thus, we will remove Trip one as this is T0 and there would be no decomposition and trip 3 because there are data missing. 

# subset 2
decomp_bags_model_data<- subset(decomp_bags, Trip == "2")

# Fit a model of percent decomposition by treatment x site, 

decomp_model<- lm(prc_Delta ~  Treatment * Site , data = decomp_bags_model_data, na.action = na.exclude)

#the test suggests the random factor is Necessary, we will not remove.

summary(decomp_model)
anova(decomp_model)

# Check residuals
plot(decomp_model)
qqnorm(resid(decomp_model)); qqline(resid(decomp_model))

# Get the estimated marginal means
emm_results <- emmeans(decomp_model, pairwise ~ Treatment | Site)

# View the results
emm_results


#to simplify this anaylysis and explore the interaction more, we weill break down the analysis by site.

# beach
decomp_bags_sub_beach<- subset(decomp_bags_model_data, Site == "Beach")
view(decomp_bags_sub_beach)

#two-sample t-test

t_test_result <- t.test(prc_Delta ~ Treatment, data = decomp_bags_sub_beach)

print(t_test_result)


#forest
decomp_bags_sub_forest <- subset(decomp_bags_model_data, Site == "Forest")
view(decomp_bags_sub_forest)

#two-sample t-test

t_test_result_f <- t.test(prc_Delta ~ Treatment, data = decomp_bags_sub_forest)

print(t_test_result_f)


# the other analysis that I'd like here is to compare Large Beach to Large Forest to compare overall decomposition

#large bags sub
decomp_bags_sub_large <- subset(decomp_bags_model_data, Treatment == "large")
view(decomp_bags_sub_large)

#two-sample t-test

t_test_result_large <- t.test(prc_Delta ~ Site, data = decomp_bags_sub_large)

print(t_test_result_large)


```


#Plant Communities 
## Interior survey

```{r}

interiorsurvey<-read.csv("interior_survey.csv", header = TRUE)

is.factor(interiorsurvey$Treatment)
interiorsurvey$Treatment<-as.factor(interiorsurvey$Treatment)
is.factor(interiorsurvey$Block)
interiorsurvey$Block<-as.factor(interiorsurvey$Block)
interiorsurvey$Trip<-as.factor(interiorsurvey$Trip)


###############################subset out the dunes############################
interior_sub_dune<-subset(interiorsurvey, Site == "Dunes")

#subset trip 1
int_sub_dune_1<-subset(interior_sub_dune, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
intdune_df_sargassum <- int_sub_dune_1[int_sub_dune_1$Treatment == "sargasso", ]
intdune_df_control <- int_sub_dune_1[int_sub_dune_1$Treatment == "control", ]

### effect sizes for trip 1
cohen.d(intdune_df_sargassum$bermuda,intdune_df_control$bermuda)
cohen.d(intdune_df_sargassum$other_plants,intdune_df_control$other_plants)


#subset trip 2
int_sub_dune_2<-subset(interior_sub_dune, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
intdune_df_sargassum2 <- int_sub_dune_2[int_sub_dune_2$Treatment == "sargasso", ]
intdune_df_control2 <- int_sub_dune_2[int_sub_dune_2$Treatment == "control", ]

### effect sizes for trip 2
cohen.d(intdune_df_sargassum2$bermuda,intdune_df_control2$bermuda)
cohen.d(intdune_df_sargassum2$other_plants,intdune_df_control2$other_plants)
cohen.d(intdune_df_sargassum2$bare_prc_cover,intdune_df_control2$bare_prc_cover)
cohen.d(intdune_df_sargassum2$leaf_prc_cover,intdune_df_control2$leaf_prc_cover)
cohen.d(intdune_df_sargassum2$other_plant_cover,intdune_df_control2$other_plant_cover)


#subset trip 3
int_sub_dune_3<-subset(interior_sub_dune, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
intdune_df_sargassum3 <- int_sub_dune_3[int_sub_dune_3$Treatment == "sargasso", ]
intdune_df_control3 <- int_sub_dune_3[int_sub_dune_3$Treatment == "control", ]

### effect sizes for trip 3
cohen.d(intdune_df_sargassum3$bermuda,intdune_df_control3$bermuda)
cohen.d(intdune_df_sargassum3$other_plants,intdune_df_control3$other_plants)
cohen.d(intdune_df_sargassum3$bare_prc_cover,intdune_df_control3$bare_prc_cover)
cohen.d(intdune_df_sargassum3$leaf_prc_cover,intdune_df_control3$leaf_prc_cover)
cohen.d(intdune_df_sargassum3$other_plant_cover,intdune_df_control3$other_plant_cover)


#subset trip 4
int_sub_dune_4<-subset(interior_sub_dune, Trip == "4")

# Subset the data for 'Sargassum' and 'Control' treatments
intdune_df_sargassum4 <- int_sub_dune_4[int_sub_dune_4$Treatment == "sargasso", ]
intdune_df_control4 <- int_sub_dune_4[int_sub_dune_4$Treatment == "control", ]

### effect sizes for trip 4
cohen.d(intdune_df_sargassum4$bermuda,intdune_df_control4$bermuda)
cohen.d(intdune_df_sargassum4$other_plants,intdune_df_control4$other_plants)
cohen.d(intdune_df_sargassum4$bare_prc_cover,intdune_df_control4$bare_prc_cover)
cohen.d(intdune_df_sargassum4$leaf_prc_cover,intdune_df_control4$leaf_prc_cover)
cohen.d(intdune_df_sargassum4$other_plant_cover,intdune_df_control4$other_plant_cover)


###########subset out the Jungle
interior_sub_Jungle<-subset(interiorsurvey, Site == "Jungle")

#subset trip 1
int_sub_Jungle_1<-subset(interior_sub_Jungle, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
intJungle_df_sargassum <- int_sub_Jungle_1[int_sub_Jungle_1$Treatment == "sargasso", ]
intJungle_df_control <- int_sub_Jungle_1[int_sub_Jungle_1$Treatment == "control", ]

### effect sizes for trip 1
cohen.d(intJungle_df_sargassum$bermuda,intJungle_df_control$bermuda)
cohen.d(intJungle_df_sargassum$total_plant_prc_cover,intJungle_df_control$total_plant_prc_cover)
cohen.d(intJungle_df_sargassum$bare_prc_cover,intJungle_df_control$bare_prc_cover)
cohen.d(intJungle_df_sargassum$leaf_prc_cover,intJungle_df_control$leaf_prc_cover)
cohen.d(intJungle_df_sargassum$other_plants,intJungle_df_control$other_plants)


#subset trip 2
int_sub_Jungle_2<-subset(interior_sub_Jungle, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
intJungle_df_sargassum2 <- int_sub_Jungle_2[int_sub_Jungle_2$Treatment == "sargasso", ]
intJungle_df_control2 <- int_sub_Jungle_2[int_sub_Jungle_2$Treatment == "control", ]

### effect sizes for trip 2
cohen.d(intJungle_df_sargassum2$bermuda,intJungle_df_control2$bermuda)
cohen.d(intJungle_df_sargassum2$total_plant_prc_cover,intJungle_df_control2$total_plant_prc_cover)
cohen.d(intJungle_df_sargassum2$bare_prc_cover,intJungle_df_control2$bare_prc_cover)
cohen.d(intJungle_df_sargassum2$leaf_prc_cover,intJungle_df_control2$leaf_prc_cover)
cohen.d(intJungle_df_sargassum2$other_plants,intJungle_df_control2$other_plants)

#subset trip 3
int_sub_Jungle_3<-subset(interior_sub_Jungle, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
intJungle_df_sargassum3 <- int_sub_Jungle_3[int_sub_Jungle_3$Treatment == "sargasso", ]
intJungle_df_control3 <- int_sub_Jungle_3[int_sub_Jungle_3$Treatment == "control", ]

### effect sizes for trip 3
cohen.d(intJungle_df_sargassum3$bermuda,intJungle_df_control3$bermuda)
cohen.d(intJungle_df_sargassum3$total_plant_prc_cover,intJungle_df_control3$total_plant_prc_cover)
cohen.d(intJungle_df_sargassum3$bare_prc_cover,intJungle_df_control3$bare_prc_cover)
cohen.d(intJungle_df_sargassum3$leaf_prc_cover,intJungle_df_control3$leaf_prc_cover)
cohen.d(intJungle_df_sargassum3$other_plants,intJungle_df_control3$other_plants)

##########the effect, while negligeble, maybe due to the cucumber plant that took over from outside in, we can test this by removing plot 1########################3

#subset trip 3
cuc_int_sub_Jungle_3<-subset(int_sub_Jungle_3, !Block == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
cuc_intJungle_df_sargassum3 <- cuc_int_sub_Jungle_3[cuc_int_sub_Jungle_3$Treatment == "sargasso", ]
cuc_intJungle_df_control3 <- cuc_int_sub_Jungle_3[cuc_int_sub_Jungle_3$Treatment == "control", ]

### effect sizes for trip 3
cohen.d(cuc_intJungle_df_sargassum3$other_plants,cuc_intJungle_df_control3$other_plants)



#subset trip 4
int_sub_Jungle_4<-subset(interior_sub_Jungle, Trip == "4")

# Subset the data for 'Sargassum' and 'Control' treatments
intJungle_df_sargassum4 <- int_sub_Jungle_4[int_sub_Jungle_4$Treatment == "sargasso", ]
intJungle_df_control4 <- int_sub_Jungle_4[int_sub_Jungle_4$Treatment == "control", ]

### effect sizes for trip 4
cohen.d(intJungle_df_sargassum4$bermuda,intJungle_df_control4$bermuda)
cohen.d(intJungle_df_sargassum4$total_plant_prc_cover,intJungle_df_control4$total_plant_prc_cover)
cohen.d(intJungle_df_sargassum4$bare_prc_cover,intJungle_df_control4$bare_prc_cover)
cohen.d(intJungle_df_sargassum4$leaf_prc_cover,intJungle_df_control4$leaf_prc_cover)
cohen.d(intJungle_df_sargassum4$other_plants,intJungle_df_control4$other_plants)


```

###percent cover
```{r}

#import dataset
interiorrsurvey<-read.csv("interior_survey.csv", header = TRUE)

interior_survey_long<- interiorsurvey %>%
  gather(key = "cover_type", value = "percent_cover", c(57,60))
interior_survey_long$Trip<-as.factor(interior_survey_long$Trip)

plant_int_data <- interior_survey_long %>%
  group_by(Trip,Site,Treatment,cover_type) %>%
  summarize(mean_cover = mean(percent_cover), se = sd(percent_cover) / sqrt(n()))
view(plant_int_data)


#graph it
interior_pc_cover<- ggplot(data = interior_survey_long,
            aes(
              x = Trip,
              y = percent_cover,
              shape = cover_type,
              color = Treatment
            )) +
  stat_summary(geom = "point", fun = mean, position = position_dodge(width = 1),
               size = 2L, stroke = 1) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 1),
               width = 0.2) +  # Adjust the width of the error bars as needed
  labs(
    x = " ",
    y = "Mean Percent Cover",
    title = "Interior Mean Percent Cover",
    shape = "Cover type"
  ) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 23")
  ) +
  facet_grid(vars(), vars(Site)) +
  theme_minimal() +
  scale_fill_manual(values = c("grey44", "black")) +
  scale_color_manual(values = c("grey44", "black")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines")
  )

interior_pc_cover


#lets look at how much sargassum additon changed grass percent cover over the entire experiment. to do this we can compare controls from august 2022 to sargassum addition august 2023 using a T test. should be near ~10% to 70%

#first subset out the beach treatment
grass_comparison<-subset(interior_survey_long, Site == "Dunes")

#now subset the grass treatment 
grass_comparison2<-subset(grass_comparison, cover_type == "bermuda")

# Subset the data for Trip 1 Control
trip1_control <- subset(grass_comparison2, Trip == "1" & Treatment == "control")

# Subset the data for Trip 4 Sargassum
trip4_sargasso <- subset(grass_comparison2, Trip == "4" & Treatment == "sargasso")

# Perform t-test
t_test_result <- t.test(trip1_control$percent_cover, trip4_sargasso$percent_cover)

# Print the result
print(t_test_result)

#now lets look at the forest
#first subset out the beach treatment
grass_comparison_F<-subset(interior_survey_long, Site == "Forest")

#now subset the grass treatment 
grass_comparison_F2<-subset(grass_comparison, cover_type == "bermuda")

# Subset the data for Trip 1 Control
trip1_control <- subset(grass_comparison2, Trip == "1" & Treatment == "control")

# Subset the data for Trip 4 Sargassum
trip4_sargasso <- subset(grass_comparison2, Trip == "4" & Treatment == "sargasso")

# Perform t-test
t_test_result <- t.test(trip1_control$percent_cover, trip4_sargasso$percent_cover)

# Print the result
print(t_test_result)





```

###effect sizes
```{r}

interioreffsize1 <-read.csv("interior_graphs.csv",header=T)
interioreffsize <-subset(interioreffsize1, Data == "eff_size")

interioreffsize<-subset(interioreffsize, !Cover == "all_plants")
interioreffsize<-subset(interioreffsize, !Cover ==  "bare" ) 
interioreffsize<-subset(interioreffsize, !Cover == "leaf_litter")
interioreffsize$Trip<-as.factor(interioreffsize$Trip)
interioreffsize$Effect_size<-as.numeric(interioreffsize$Effect_size)


int_eff_plot <- ggplot(interioreffsize, aes(x = Trip, y = Effect_size, shape = Cover)) +
  labs(y = "Effect size", title = "") +
  geom_path(aes(group = Cover), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar.23", "4" = "Aug.23")
  ) +
    scale_shape_discrete(labels = c("grass", "other plants")) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    strip.text = element_text(size = 12)
  ) +
  xlab("") +
    theme_minimal()  +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"))

int_eff_plot

```

## Perimeter Survey
```{r}

perimetersurvey<-read.csv("perimeter_survey.csv", header = TRUE)

perimetersurvey$Treatment<-as.factor(perimetersurvey$Treatment)
perimetersurvey$Block<-as.factor(perimetersurvey$Block)
perimetersurvey$Trip<-as.factor(perimetersurvey$Trip)

###########subset out the dunes
perimeter_sub_dune<-subset(perimetersurvey, Site == "Dunes")

#subset trip 1
per_sub_dune_1<-subset(perimeter_sub_dune, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
perdune_df_sargassum <- per_sub_dune_1[per_sub_dune_1$Treatment == "sargasso", ]
perdune_df_control <- per_sub_dune_1[per_sub_dune_1$Treatment == "control", ]

### effect sizes for trip 1
cohen.d(perdune_df_sargassum$bermuda,perdune_df_control$bermuda)
cohen.d(perdune_df_sargassum$other_plants,perdune_df_control$other_plants)
cohen.d(perdune_df_sargassum$total_plant_prc_cover,perdune_df_control$total_plant_prc_cover)
cohen.d(perdune_df_sargassum$bare_prc_cover,perdune_df_control$bare_prc_cover)
cohen.d(perdune_df_sargassum$dead_prc_cover,perdune_df_control$dead_prc_cover)



#subset trip 2
per_sub_dune_2<-subset(perimeter_sub_dune, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
perdune_df_sargassum2 <- per_sub_dune_2[per_sub_dune_2$Treatment == "sargasso", ]
perdune_df_control2 <- per_sub_dune_2[per_sub_dune_2$Treatment == "control", ]

### effect sizes for trip 2
cohen.d(perdune_df_sargassum2$bermuda,perdune_df_control2$bermuda)
cohen.d(perdune_df_sargassum2$other_plants,perdune_df_control2$other_plants)
cohen.d(perdune_df_sargassum2$total_plant_prc_cover,perdune_df_control2$total_plant_prc_cover)
cohen.d(perdune_df_sargassum2$bare_prc_cover,perdune_df_control2$bare_prc_cover)
cohen.d(perdune_df_sargassum2$dead_prc_cover,perdune_df_control2$dead_prc_cover)



#subset trip 3
per_sub_dune_3<-subset(perimeter_sub_dune, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
perdune_df_sargassum3 <- per_sub_dune_3[per_sub_dune_3$Treatment == "sargasso", ]
perdune_df_control3 <- per_sub_dune_3[per_sub_dune_3$Treatment == "control", ]

### effect sizes for trip 3
cohen.d(perdune_df_sargassum3$bermuda,perdune_df_control3$bermuda)
cohen.d(perdune_df_sargassum3$other_plants,perdune_df_control3$other_plants)
cohen.d(perdune_df_sargassum3$total_plant_prc_cover,perdune_df_control3$total_plant_prc_cover)
cohen.d(perdune_df_sargassum3$bare_prc_cover,perdune_df_control3$bare_prc_cover)
cohen.d(perdune_df_sargassum3$dead_prc_cover,perdune_df_control3$dead_prc_cover)


#subset trip 4
per_sub_dune_4<-subset(perimeter_sub_dune, Trip == "4")

# Subset the data for 'Sargassum' and 'Control' treatments
perdune_df_sargassum4 <- per_sub_dune_4[per_sub_dune_4$Treatment == "sargasso", ]
perdune_df_control4 <- per_sub_dune_4[per_sub_dune_4$Treatment == "control", ]

### effect sizes for trip 4
cohen.d(perdune_df_sargassum4$bermuda,perdune_df_control4$bermuda)
cohen.d(perdune_df_sargassum$other_plants,perdune_df_control4$other_plants)
cohen.d(perdune_df_sargassum4$total_plant_prc_cover,perdune_df_control4$total_plant_prc_cover)
cohen.d(perdune_df_sargassum4$bare_prc_cover,perdune_df_control4$bare_prc_cover)
cohen.d(perdune_df_sargassum4$dead_prc_cover,perdune_df_control4$dead_prc_cover)


###########subset out the Jungle
perimeter_sub_Jungle<-subset(perimetersurvey, Site == "Jungle")

#subset trip 1
per_sub_Jungle_1<-subset(perimeter_sub_Jungle, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
perJungle_df_sargassum <- per_sub_Jungle_1[per_sub_Jungle_1$Treatment == "sargasso", ]
perJungle_df_control <- per_sub_Jungle_1[per_sub_Jungle_1$Treatment == "control", ]

### effect sizes for trip 1
cohen.d(perJungle_df_sargassum$bermuda,perJungle_df_control$bermuda)
cohen.d(perJungle_df_sargassum$other_plants,perJungle_df_control$other_plants)
cohen.d(perJungle_df_sargassum$total_plant_prc_cover,perJungle_df_control$total_plant_prc_cover)
cohen.d(perJungle_df_sargassum$bare_prc_cover,perJungle_df_control$bare_prc_cover)
cohen.d(perJungle_df_sargassum$dead_prc_cover,perJungle_df_control$dead_prc_cover)


#subset trip 2
per_sub_Jungle_2<-subset(perimeter_sub_Jungle, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
perJungle_df_sargassum2 <- per_sub_Jungle_2[per_sub_Jungle_2$Treatment == "sargasso", ]
perJungle_df_control2 <- per_sub_Jungle_2[per_sub_Jungle_2$Treatment == "control", ]

### effect sizes for trip 2
cohen.d(perJungle_df_sargassum2$bermuda,perJungle_df_control2$bermuda)
cohen.d(perJungle_df_sargassum2$other_plants,perJungle_df_control2$other_plants)
cohen.d(perJungle_df_sargassum2$total_plant_prc_cover,perJungle_df_control2$total_plant_prc_cover)
cohen.d(perJungle_df_sargassum2$bare_prc_cover,perJungle_df_control2$bare_prc_cover)
cohen.d(perJungle_df_sargassum2$dead_prc_cover,perJungle_df_control2$dead_prc_cover)


#subset trip 3
per_sub_Jungle_3<-subset(perimeter_sub_Jungle, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
perJungle_df_sargassum3 <- per_sub_Jungle_3[per_sub_Jungle_3$Treatment == "sargasso", ]
perJungle_df_control3 <- per_sub_Jungle_3[per_sub_Jungle_3$Treatment == "control", ]

### effect sizes for trip 3
cohen.d(perJungle_df_sargassum3$bermuda,perJungle_df_control3$bermuda)
cohen.d(perJungle_df_sargassum3$other_plants,perJungle_df_control3$other_plants)
cohen.d(perJungle_df_sargassum3$total_plant_prc_cover,perJungle_df_control3$total_plant_prc_cover)
cohen.d(perJungle_df_sargassum3$bare_prc_cover,perJungle_df_control3$bare_prc_cover)
cohen.d(perJungle_df_sargassum3$dead_prc_cover,perJungle_df_control3$dead_prc_cover)


#subset trip 4
per_sub_Jungle_4<-subset(perimeter_sub_Jungle, Trip == "4")

# Subset the data for 'Sargassum' and 'Control' treatments
perJungle_df_sargassum4 <- per_sub_Jungle_4[per_sub_Jungle_4$Treatment == "sargasso", ]
perJungle_df_control4 <- per_sub_Jungle_4[per_sub_Jungle_4$Treatment == "control", ]

### effect sizes for trip 4
cohen.d(perJungle_df_sargassum4$bermuda,perJungle_df_control4$bermuda)
cohen.d(perJungle_df_sargassum4$other_plants,perJungle_df_control4$other_plants)
cohen.d(perJungle_df_sargassum4$total_plant_prc_cover,perJungle_df_control4$total_plant_prc_cover)
cohen.d(perJungle_df_sargassum4$bare_prc_cover,perJungle_df_control4$bare_prc_cover)
cohen.d(perJungle_df_sargassum4$dead_prc_cover,perJungle_df_control4$dead_prc_cover)



```



###percent cover
```{r}

#import dataset
perimetersurvey<-read.csv("perimeter_survey.csv", header = TRUE)

perimeter_survey_long<- perimetersurvey %>%
  gather(key = "cover_type", value = "percent_cover", c(60,61))
perimeter_survey_long$Trip<-as.factor(perimeter_survey_long$Trip)


plant_per_data <- perimeter_survey_long %>%
  group_by(Trip,Site,Treatment,cover_type) %>%
  summarize(mean_cover = mean(percent_cover), se = sd(percent_cover) / sqrt(n()))
view(plant_per_data)

#graph it


perimeter_pc_cover<- ggplot(data = perimeter_survey_long,
            aes(
              x = Trip,
              y = percent_cover,
              shape = cover_type,
              color = Treatment
            )) +
  stat_summary(geom = "point", fun = mean, position = position_dodge(width = 1),
               size = 2L, stroke = 1) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 1),
               width = 0.2) +  # error bars
  labs(
    x = " ",
    y = "Mean Percent Cover",
    title = "perimeter Mean Percent Cover",
    shape = "Cover type"
  ) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 23")
  ) +
  facet_grid(vars(), vars(Site)) +
  theme_minimal() +
  scale_fill_manual(values = c("grey44", "black")) +
  scale_color_manual(values = c("grey44", "black")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines")
  )

perimeter_pc_cover


```


###effect sizes
```{r}

perimetereffsize1 <-read.csv("perimeter_graphs.csv",header=T)
perimetereffsize <-subset(perimetereffsize1, Data == "eff_size")

perimetereffsize<-subset(perimetereffsize, !Cover == "all_plants")
perimetereffsize<-subset(perimetereffsize, !Cover ==  "dead" ) 
perimetereffsize<-subset(perimetereffsize, !Cover ==  "bare" ) 
perimetereffsize$Trip<-as.factor(perimetereffsize$Trip)
perimetereffsize$Effect_size<-as.numeric(perimetereffsize$Effect_size)


per_eff_plot <- ggplot(perimetereffsize, aes(x = Trip, y = Effect_size, shape = Cover)) +
  labs(y = "Effect size", title = "") +
  geom_path(aes(group = Cover), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar.23", "4" = "Aug.23")
  ) +
    scale_shape_discrete(labels = c("grass", "other plants")) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 12),
    plot.title = element_text(size = 12),
    axis.text.x = element_text(size = 12),
    strip.text = element_text(size = 12)
  ) +
  xlab("") +
    theme_minimal()  +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"))

per_eff_plot


```

##Ladder
```{r}
ladder<-read.csv("ladder.csv", header = TRUE)
ladder$Trip<-as.factor(ladder$Trip)
ladder$Quadrat<-as.factor(ladder$Quadrat)

#make data long 
laddersurvey_long <- gather(ladder, key = "cover_type", value = "percent_cover", 10:11, -c(1:5),) 
#use this for graphing purposes
laddersurvey_long$percent_cover<-as.numeric(laddersurvey_long$percent_cover)
laddersurvey_long$Trip<-as.factor(laddersurvey_long$Trip)
laddersurvey_long$Quadrat <- as.factor(laddersurvey_long$Quadrat)

#graph it
#lets look at the distance effect of sargassum on percent cover of grass and all other plants in the jungle 


ladder<- laddersurvey_long %>%
  ggplot() +
  aes(x = Dist_from_edge, y = percent_cover, fill = cover_type) +
  stat_summary(
    fun.y = "mean",
    geom = "bar",
    position = "dodge",
    width = 0.7
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    position = position_dodge(width = 0.7),
    width = 0.25
  ) +
  scale_fill_hue(direction = 1) +
  labs(
    x = "Distance from plot (m)",
    y = "Percent cover",
    fill = "Cover type") +
  theme_minimal() +
  facet_grid(vars(Trip), vars(Site), scales = "free")

ladder

 

laddersurvey_long %>%
filter(Quadrat != "0") %>%
  ggplot(aes(x = Quadrat, fill = Treatment, y = percent_cover)) +
  geom_bar(position = "dodge", stat = "identity") +
  stat_summary(
    fun.data = "mean_se",
    geom = "errorbar",
    position = position_dodge(width = 0.7),
    width = 0.25
  ) +
  theme_minimal() +
  facet_grid(vars(Trip), vars(Site))





#####now lets look at the two sites seperately to see if there is a spillover effect away from the prot perimeter (D1 and D2 distances are not equal, thus, we cannot use this for the paper, just want to know if there is some distance effect)

# Convert Quadrat to a factor to ensure correct grouping
ladder$Quadrat <- as.factor(ladder$Quadrat)
# remove the "0" treatment since those data are used toward the perimieter surveys and we are now interested in seeing if there is a treatment effect BEYOND those

laddersurvey_dist<-subset(ladder, !Quadrat == "0")


ladder_beach<-subset(laddersurvey_dist, Site == "Beach")

#subset trips. No use in looking at trip one because there is no way sargasso dumping affects plant cover beyond the pile instantly
ladder_beach_2<-subset(ladder_beach, Trip == "2")
###now at Distance 2
ladder_beach_2_D1<-subset(ladder_beach_2, Dist_from_edge == "Distance 1")

# Subset the data for 'Sargassum' and 'Control' treatments
ladder_beach_2_D1_sarg <- ladder_beach_2_D1[ladder_beach_2_D1$Treatment == "Sargassum", ]
ladder_beach_2_D1_con <- ladder_beach_2_D1[ladder_beach_2_D1$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(ladder_beach_2_D1_sarg$Bermuda,ladder_beach_2_D1_con$Bermuda)
cohen.d(ladder_beach_2_D1_sarg$Other_pants,ladder_beach_2_D1_con$Other_pants)

#effect size for Grass does not exist because there is zeros in both treatments, further, the effect for other plants is negligeble. 


### to be on the safe side, lets lets look at this effect (or lack thereof) for trip 3
ladder_beach_3<-subset(ladder_beach, Trip == "3")
###now at Distance 2
ladder_beach_3_D1<-subset(ladder_beach_3, Dist_from_edge == "Distance 1")

# Subset the data for 'Sargassum' and 'Control' treatments
ladder_beach_3_D1_sarg <- ladder_beach_3_D1[ladder_beach_3_D1$Treatment == "Sargassum", ]
ladder_beach_3_D1_con <- ladder_beach_3_D1[ladder_beach_3_D1$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(ladder_beach_3_D1_sarg$Bermuda,ladder_beach_3_D1_con$Bermuda)
cohen.d(ladder_beach_3_D1_sarg$Other_pants,ladder_beach_3_D1_con$Other_pants)
#effect size for Grass does not exist because there is zeros in both treatments, further, the effect for other plants is negligeble. 



ladder_forest<-subset(laddersurvey_dist, Site == "Forest")

```

#Arthropods
##pitfalls

```{r}

#load data
pitfalls <-read.csv("Pitfalls.csv",header=T)
pitfalls_long<- pitfalls %>%
  gather(key = "Taxa", value = "Count", c(8,13,15))
pitfalls_long$Trip<-as.factor(pitfalls_long$Trip)


#fix names and calculate Mean and SE

pitfalls_data <- pitfalls_long %>%
  mutate(
    Site = ifelse(Site == "Dune", "Beach", ifelse(Site == "Jungle", "Forest", Site)),
    Trip = case_when(
      Trip == 1 ~ "Aug. 22",
      Trip == 2 ~ "Nov. 22",
      Trip == 3 ~ "Mar. 23",
      TRUE ~ as.character(Trip)  
    ),
    Treatment = case_when(
      Treatment == "Control" ~ "control",
      Treatment == "Sargassum" ~ "sargasso",
      TRUE ~ as.character(Treatment)  
    )
  ) %>%
  group_by(Site, Trip, Treatment, Taxa) %>%
  summarize(
    Mean = round(mean(Count), digits = 2), 
    SE = round(sd(Count) / sqrt(n()), digits = 2)  
  )

view(pitfalls_data)

#make data wide by "Site"
pit_wide_data <- pitfalls_data %>%
  pivot_wider(
    id_cols = c("Trip", "Treatment", "Taxa"),
    names_from = Site,
    values_from = c(Mean, SE),
    names_glue = "{Site}_{.value}")

View(pit_wide_data)


#make wider by Trip
pit_wide_data2 <- pit_wide_data %>%
  pivot_wider(
    id_cols = c( "Treatment", "Taxa"),
    names_from = Trip,
    values_from = c(Beach_Mean, Beach_SE,
    Forest_Mean, Forest_SE),
    names_glue = "{Trip}_{.value}" )

view(pit_wide_data2)

#make sure Means are next to related SEs
pit_wide_data3 <- pit_wide_data2 %>%
  select(
    # List the columns in order
    "Taxa", "Treatment",
    "Aug. 22_Beach_Mean", "Aug. 22_Beach_SE",
    "Nov. 22_Beach_Mean", "Nov. 22_Beach_SE",
    "Mar. 23_Beach_Mean", "Mar. 23_Beach_SE",
    "Aug. 22_Forest_Mean", "Aug. 22_Forest_SE",
    "Nov. 22_Forest_Mean", "Nov. 22_Forest_SE",
    "Mar. 23_Forest_Mean", "Mar. 23_Forest_SE",)

View(pit_wide_data3)

# Rename the orders in Taxa
pit_wide_data3 <- pit_wide_data3 %>%
  mutate(
    Taxa = case_when(
      Taxa == "Amphipods" ~ "Amphipoda",
      Taxa == "Arachnid" ~ "Arachnida",
      Taxa == "Hymenoptera" ~ "Hymenoptera",
      TRUE ~ Taxa  # Keep other values as is
    )
  )
#order by Taxa
pit_wide_data3 <- pit_wide_data3 %>%
  arrange(Taxa)

#Rename columns to just show Mean and SE
### !!!!important!!!! notice i staggered spaces in the name so that i can use the same column names for all Means and SEs
pit_wide_data4 <- pit_wide_data3 %>%
  rename(
     "Mean" = "Aug. 22_Beach_Mean", "SE   " = "Aug. 22_Beach_SE",
     " Mean" = "Nov. 22_Beach_Mean", "SE  " = "Nov. 22_Beach_SE",
     "  Mean" = "Mar. 23_Beach_Mean", "SE " = "Mar. 23_Beach_SE",
     "Mean " = "Aug. 22_Forest_Mean", "  SE"  = "Aug. 22_Forest_SE",
     "Mean  " = "Nov. 22_Forest_Mean", " SE" = "Nov. 22_Forest_SE",
     "Mean   " = "Mar. 23_Forest_Mean", "SE" = "Mar. 23_Forest_SE")


kbl(pit_wide_data4) %>%
  kable_classic("striped", full_width = T)%>%
  add_header_above(c(" " = 2, "Aug. 22" = 2, "Nov. 22" = 2, "Mar. 23" = 2, "Aug. 22" = 2, "Nov. 22" = 2, "Mar. 23" = 2)) %>%
  add_header_above(c(" " = 2, "Beach" = 6, "Forest" = 6)) 



```

####pitfalls analysis


```{r}

stickiesmod1 <-read.csv("stickies.csv",header=T)

############################################dunes############################################
#subset out the dunes
eff_sub_dune<-subset(stickiesmod1, Site == "Dune")

#subset trip 1
eff_sub_dune_1<-subset(eff_sub_dune, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
dune_df_sargassum <- eff_sub_dune_1[eff_sub_dune_1$Treatment == "Sargassum", ]
dune_df_control <- eff_sub_dune_1[eff_sub_dune_1$Treatment == "Control", ]

### effect sizes for trip 1
cohen.d(dune_df_sargassum$Amphipods,dune_df_control$Amphipods)
cohen.d(dune_df_sargassum$Diptera,dune_df_control$Diptera)
cohen.d(dune_df_sargassum$Hymenoptera,dune_df_control$Hymenoptera)

#subset trip 2
eff_sub_dune_2<-subset(eff_sub_dune, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
dune_df_sargassum2 <- eff_sub_dune_2[eff_sub_dune_2$Treatment == "Sargassum", ]
dune_df_control2 <- eff_sub_dune_2[eff_sub_dune_2$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(dune_df_sargassum2$Amphipods,dune_df_control2$Amphipods)
cohen.d(dune_df_sargassum2$Diptera,dune_df_control2$Diptera)
cohen.d(dune_df_sargassum2$Hymenoptera,dune_df_control2$Hymenoptera)

#subset trip 3
eff_sub_dune_3<-subset(eff_sub_dune, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
dune_df_sargassum3 <- eff_sub_dune_3[eff_sub_dune_3$Treatment == "Sargassum", ]
dune_df_control3 <- eff_sub_dune_3[eff_sub_dune_3$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(dune_df_sargassum3$Amphipods,dune_df_control3$Amphipods)
cohen.d(dune_df_sargassum3$Diptera,dune_df_control3$Diptera)
cohen.d(dune_df_sargassum3$Hymenoptera,dune_df_control3$Hymenoptera)


############################################jungle############################################
#subset out the dunes
eff_sub_jungle<-subset(stickiesmod1, Site == "Jungle")

#subset trip 1
eff_sub_jungle_1<-subset(eff_sub_jungle, Trip == "1")

# Subset the data for 'Sargassum' and 'Control' treatments
jungle_df_sargassum <- eff_sub_jungle_1[eff_sub_jungle_1$Treatment == "Sargassum", ]
jungle_df_control <- eff_sub_jungle_1[eff_sub_jungle_1$Treatment == "Control", ]

### effect sizes for trip 1
cohen.d(jungle_df_sargassum$Amphipods,jungle_df_control$Amphipods)
cohen.d(jungle_df_sargassum$Diptera,jungle_df_control$Diptera)
cohen.d(jungle_df_sargassum$Hymenoptera,jungle_df_control$Hymenoptera)

#subset trip 2
eff_sub_jungle_2<-subset(eff_sub_jungle, Trip == "2")

# Subset the data for 'Sargassum' and 'Control' treatments
jungle_df_sargassum2 <- eff_sub_jungle_2[eff_sub_jungle_2$Treatment == "Sargassum", ]
jungle_df_control2 <- eff_sub_jungle_2[eff_sub_jungle_2$Treatment == "Control", ]

### effect sizes for trip 2
cohen.d(jungle_df_sargassum2$Amphipods,jungle_df_control2$Amphipods)
cohen.d(jungle_df_sargassum2$Diptera,jungle_df_control2$Diptera)
cohen.d(jungle_df_sargassum2$Hymenoptera,jungle_df_control2$Hymenoptera)

#subset trip 3
eff_sub_jungle_3<-subset(eff_sub_jungle, Trip == "3")

# Subset the data for 'Sargassum' and 'Control' treatments
jungle_df_sargassum3 <- eff_sub_jungle_3[eff_sub_jungle_3$Treatment == "Sargassum", ]
jungle_df_control3 <- eff_sub_jungle_3[eff_sub_jungle_3$Treatment == "Control", ]

### effect sizes for trip 3
cohen.d(jungle_df_sargassum3$Amphipods,jungle_df_control3$Amphipods)
cohen.d(jungle_df_sargassum3$Diptera,jungle_df_control3$Diptera)
cohen.d(jungle_df_sargassum3$Hymenoptera,jungle_df_control3$Hymenoptera)

#####Effect sizes Graph#####
#I manually input the effect sizes, lower and upper limits to a file called stickies_effsize.csv
stickieseffsize <-read.csv("sticky_effsize.csv",header=T)

stickieseffsize$Trip<-as.factor(stickieseffsize$Trip)



sticky_eff_plot <- ggplot(stickieseffsize, aes(x = Trip, y = Effect_size, shape = Order)) +
  labs(y = "Effect size", shape = "Taxa", title = "Aerial traps") +
  geom_path(aes(group = Order), position = position_dodge(0.5), color = "dark gray") +
  theme_bw() +
  facet_wrap(vars(Site), ncol = 2L, labeller = as_labeller(c("Dune" = "Beach", "Jungle" = "Forest"))) +
     geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
    scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "", "2" = "", "3" = "") ) +
  theme(
    axis.text = element_text(size=12),
    axis.title = element_text(size=12),
    plot.title = element_text(size=12),
    axis.text.x = element_text(size=12),
    strip.text = element_text(size=12)) +
   xlab("") 
#geom_pointrange(aes(ymin = lower_ci, ymax = upper_ci), position = position_dodge(0.5), color = " gray") +

sticky_eff_plot


```



#Chemicals and gases
## quick nutrients

```{r}

##quick nutrients

chemicals<-read.csv("Soil_Chemicals2.csv",header=T)

#subset experiment
SM_chem_sub<-subset(chemicals, Exp == "Sargasso Manipulation")
SG_chem<-subset(SM_chem_sub, Type == "Sargassum")
pile_chem<-subset(SG_chem, Trt == "Sediment")
pile_chem$Block<-as.factor(pile_chem$Block)
pile_chem$Trip<-as.factor(pile_chem$Trip)
pile_chem$pH<-as.numeric(pile_chem$pH)
pile_chem$KH<-as.numeric(pile_chem$KH)


#explore
#nitrate by block at beach and forest for 4 trips
ggplot(pile_chem) +
  aes(x = Site, y = Nitrate, fill = Block) +
  geom_boxplot() +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_grid(vars(Trip), vars())




#sargasso treatment. It is way too convoluted and impossible t read this way
pile_chem %>%
 filter(Type %in% "Sargassum") %>%
 ggplot() +
 aes(x = Site, y = Nitrate, fill = Trt, colour = Block) +
 geom_boxplot() +
 scale_fill_hue(direction = 1) +
 scale_color_hue(direction = 1) +
 theme_minimal() +
 facet_grid(vars(Trip), vars())


# We are interested in the sediment nutrients rather than sediment and seaweed

pile_chem %>%
  filter(Type %in% "Sargassum") %>%
  filter(Trt %in% "Sediment") %>%
  group_by(Trip, Site) %>%
  summarize(mean_Nitrate = mean(Nitrate), se = sd(Nitrate) / sqrt(n())) %>%
  ggplot() +
  aes(x = Trip, y = mean_Nitrate, fill = Site) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean_Nitrate - se, ymax = mean_Nitrate + se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  scale_fill_hue(direction = 1) +
    scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 2024") ) +
  labs(
    x = " ",
    y = "Nitrate content (mg/L)",
    title = "",
    fill = "Sample"
  ) +
  theme_minimal()  +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"))


```

##full chem
```{r}
chemistry<-read.csv("sargasso_nitrates.csv",header=T)

chemistry$block<-as.factor(chemistry$block)
chemistry$trip<-as.factor(chemistry$trip)
chemistry$pair<-paste(sep = "",chemistry$site,chemistry$block)

view(chemistry)

hist(chemistry$NH4, 
     main = "Histogram of Variable", 
     xlab = "Variable Values",
     breaks = seq(min(chemistry$NH4), max(chemistry$NH4) + 1, by = 1))



#data are positively skewed, lets log transform
chemistry$NH4_transformed <- log1p(chemistry$NH4)

hist(chemistry$NH4_transformed , 
     main = "Histogram of Variable", 
     xlab = "Variable Values",
     breaks = seq(min(chemistry$NH4_transformed ), max(chemistry$NH4_transformed ) + 1, by = 1))
# Create a Q-Q plot to assess normality
qqnorm(chemistry$NH4_transformed )
qqline(chemistry$NH4_transformed )

#################ammonium
#make pairs column
ammonium<-lmer(rank(NH4_transformed) ~ treatment * site * trip + (1|pair), data = chemistry)

summary(ammonium)
anova(ammonium)

#look at residuals 
plot(ammonium, which = c(1, 3)) 
#looks fine

# Create emmeans object
emmeans_NH4 <- emmeans(ammonium, ~treatment + site*trip)

# Display pairwise comparisons
pairwise_NH4 <- pairs(emmeans_NH4)
print(pairwise_NH4)



##########################3nitrate

hist(chemistry$N, 
     main = "Histogram of Variable", 
     xlab = "Variable Values",
     breaks = seq(min(chemistry$N), max(chemistry$N) + 1, by = 1))

# Create a Q-Q plot to assess normality
qqnorm(chemistry$N)
qqline(chemistry$N)
#data look normal enough 

N<-lmer(rank(N) ~ treatment * site * trip + (1|pair), data = chemistry)

summary(N)
anova(N)

#look at residuals 
plot(N, which = c(1, 3)) 
#looks fine

# Create emmeans object
emmeans_N <- emmeans(N, ~treatment + site*trip)

# Display pairwise comparisons
pairwise_N <- pairs(emmeans_N)
print(pairwise_N)



#########################################DOC

hist(chemistry$C, 
     main = "Histogram of Variable", 
     xlab = "Variable Values",
     breaks = seq(min(chemistry$C), max(chemistry$C) + 1, by = 1))

# Create a Q-Q plot to assess normality
qqnorm(chemistry$C)
qqline(chemistry$C)

#data look normal enough 
C<-lm(C ~ treatment + site * trip, data = chemistry)
summary(C)
anova(C)

#look at residuals 
plot(C, which = c(1, 3)) 
#looks fine

# Create emmeans object
emmeans_C <- emmeans(C, ~treatment + site*trip)

# Display pairwise comparisons
pairwise_C <- pairs(emmeans_C)
print(pairwise_C)


```


##gasses
```{r}
CO2<-read.csv("gas_readings.csv",header=T)

CO2$plotnum<-as.factor(CO2$plotnum)
is.factor(CO2$plotnum)

view(CO2)

#lets look at the data 
hist(CO2$gCO2_m2_h)


#let’s find out if theres the overall treatment effect (sargassum addition) on soil respiration. Here we will look to see if there is an interaction between our 3 factors, thus, we will make a three way anova

#model
co2model <- lmer(rank(gCO2_m2_h) ~ site*trt*month + (1|plot) , data = CO2)
anova(co2model)
anova (lmer(rank(gCO2_m2_h) ~ site*trt*month + (1|plotnum), CO2))

summary(co2model)
# Extract residuals from the model
residuals <- residuals(co2model)

plot(predict(co2model), residuals)
abline(h = 0, col = "red")  # red line

# narmality of residuals. 
qqnorm(residuals)
qqline(residuals)
# not great but expected for ecological data

# Create emmeans object
emmeans_Co2 <- emmeans(co2model, ~treatment* site*trip)

# Display pairwise comparisons
pairwise_Co2 <- pairs(emmeans_Co2)
print(pairwise_Co2)



#all factors show significance. J. Long is specifically interested in the treatment effect and that is significant P<0.01. There is no 3 way interaction but site and trip interact. So let’s look at them separately


##################beach####################################################

CO2_B <- subset(CO2, site == "Beach")

# Subset the data for each time point
trip2_data <- subset(CO2_B, trip == "2")
trip3_data <- subset(CO2_B, trip == 3)
trip4_data <- subset(CO2_B, trip == 4)


# Mixed-effects model for trip 2
model_trip2 <- lm(co2 ~ treatment  , data = trip2_data)

# Mixed-effects model for trip 2
model_trip3 <- lm(co2 ~ treatment , data = trip3_data)

# Mixed-effects model for trip 3
model_trip4 <- lm(co2 ~ treatment , data = trip4_data)

summary(model_trip2)
summary(model_trip3)
summary(model_trip4)

##################forest####################################################

CO2_F <- subset(CO2, site == "Forest")

# Subset the data for each time point
trip2_data_F <- subset(CO2_F, trip == "2")
trip3_data_F <- subset(CO2_F, trip == 3)
trip4_data_F <- subset(CO2_F, trip == 4)


# Mixed-effects model for trip 2
F_model_trip2 <- lmer(co2 ~ treatment  + (1|rep), data = trip2_data_F)

# Mixed-effects model for trip 2
F_model_trip3 <- lmer(co2 ~ treatment + (1|rep), data = trip3_data_F)

# Mixed-effects model for trip 3
F_model_trip4 <- lmer(co2 ~ treatment + (1|rep), data = trip4_data_F)

summary(F_model_trip2)
summary(F_model_trip3)
summary(F_model_trip4)





```
#Manuscript figures

##decomp bags
````{r}
decomp_bags<-read.csv("mesh_bag_exp.csv",header = T )
decomp_bags$Rep_no.<-as.factor(decomp_bags$Rep_no.)
decomp_bags$Trip<-as.factor(decomp_bags$Trip)

view(decomp_bags)


#quick view counts of percent
ggplot(decomp_bags) +
 aes(x = Trip, fill = Treatment, weight = prc_means) +
 geom_bar(position = "dodge") +
 scale_fill_hue(direction = 1) +
 theme_minimal() +
 facet_grid(vars(Site), vars())
##############################decomposer experiment##################################
#now lets look at mean values and SE
#subset out trip 1 since its only going to show initial mass which is 100%


decomp_bags<- decomp_bags %>%
  filter(!(Trip %in% "1")) %>%
  ggplot() +
  scale_fill_manual(name = "Mesh size", 
                    values = c("large" = "light gray", "small" = "white")) +  
  scale_color_manual(name = "Mesh size",
                     values = c("large" = "black", "small" = "black")) +  
  aes(x = Site, fill = Treatment, weight = prc_means) + 
  geom_bar(position = "dodge", color = "black") +   
  geom_errorbar(aes(ymin = prc_means - prc_SE, ymax = prc_means + prc_SE), 
                position = position_dodge(width = 0.9), 
                width = 0.25) +  
  labs(y = "Percent of initial dry biomass", x = "") +
  theme_classic() +
  facet_grid(Trip ~ ., labeller = labeller(Trip = function(x) ifelse(x == "2", "Nov. 22", ifelse(x == "3", "Mar. 23", x)))) + 
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    axis.title = element_text(size = 16, color = "black"),   
    axis.text = element_text(size = 14, color = "black"),     
    axis.text.x = element_text(size = 14, color = "black"),   
    plot.title = element_text(size = 18, color = "black"),   
    legend.title = element_text(size = 16, color = "black"), 
    legend.text = element_text(size = 14, color = "black"),
    legend.position = "bottom",
    strip.text = element_text(size = 16, color = "black"),  
    plot.background = element_rect(fill = "white")) 

print(decomp_bags)

ggsave("Figure 1.jpg", plot = decomp_bags, width = 8, height = 6, units = "in", dpi = 600)



decomp_bags %>%
  filter(!(Trip %in% "1")) %>%
  ggplot() +
  scale_fill_manual(
    name = "Mesh size",
    values = c("large" = "light gray", "small" = "white"),
    labels = c("large" = "Large mesh", "small" = "Small mesh")
  ) +
  scale_color_manual(
    name = "Mesh size",
    values = c("large" = "black", "small" = "black"),
    labels = c("large" = "Large mesh", "small" = "Small mesh")
  ) +
  aes(x = Site, fill = Treatment, weight = prc_means) +
  geom_bar(position = "dodge", color = "black") +
  geom_errorbar(
    aes(ymin = prc_means - prc_SE, ymax = prc_means + prc_SE),
    position = position_dodge(width = 0.9),
    width = 0.25
  ) +
  labs(y = "Decomposition (% of initial dry biomass)", x = "") +
  theme_classic() +
  facet_grid(
    Trip ~ .,
    labeller = labeller(Trip = function(x) ifelse(x == "2", "Nov. 22", ifelse(x == "3", "Mar. 23", x)))
  ) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    axis.title = element_text(size = 16, color = "black"),
    axis.text = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    plot.title = element_text(size = 18, color = "black"),
    legend.title = element_text(size = 16, color = "black"),
    legend.text = element_text(size = 14, color = "black"),
    legend.position = "bottom",
    strip.text = element_text(size = 16, color = "black"),
    plot.background = element_rect(fill = "white")
  )

````

##decomp inverts
```{r}
decomp<-read.csv("mesh_bag_inverts.csv",header = T )
decomp$Rep_no.<-as.factor(decomp$Rep_no.)
decomp$Trip<-as.factor(decomp$Trip)

view(decomp)

decomp_inverts<- decomp %>%
  gather(key = "inverts", value = "count", c(5,6,11))

#remove small bags 
decomp_inverts2<-subset(decomp_inverts, Treatment == "large")


invert_plot <- ggplot(data = decomp_inverts2,
                      aes(
                        x = inverts,
                        y = count,
                      )) +
  stat_summary(geom = "bar", fun = mean, position = position_dodge(width = 1),
               size = 2L, stroke = 1) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 1),
               width = 0.2) +  
  theme_classic() +
  facet_grid(vars(Trip), vars(Site)) +
   theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14, color = "black"),
    axis.title = element_text(size = 16, color = "black"),
    axis.text = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    plot.title = element_text(size = 12, color = "black"),
    legend.title = element_text(size = 16, color = "black"),
    legend.text = element_text(size = 14, color = "black"),
    strip.text = element_text(size = 14, color = "black")) +  
  labs(y = "Arthroopod counts", x = "", title = "")

invert_plot

ggsave("Figure 2.jpg", plot = invert_plot, width = 8, height = 7, units = "in", dpi = 600)

```

##chemistry
```{r}



#ammonium
ammonium<-read.csv("sargasso_nitrates.csv",header=T)

ammonium$block<-as.factor(ammonium$block)
ammonium$trip<-as.factor(ammonium$trip)

NH4<-ammonium %>%
 filter(!(site %in% "")) %>%
 filter(!(treatment %in% "")) %>%
  group_by(trip, site, treatment) %>%
  summarize(mean_NH4 = mean(NH4), NH4_se = sd(NH4) / sqrt(n())) %>%
  ggplot() +
  aes(x =trip, y = mean_NH4, fill = treatment) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean_NH4 - NH4_se, ymax = mean_NH4 + NH4_se), 
                position = position_dodge(width = 0.9), width = 0.25) +
   scale_fill_manual(values = c("lightgray", "dark gray")) +
    scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 2023") ) +
  labs(
    x = " ",
    y = expression("Ammonium (mg) sediment (kg"^{-1} * ")"),
    title = "",
    fill = "Treatment") +
  theme_classic()  +
   theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
   axis.title = element_text(size = 14, color = "black"),   
    axis.text = element_text(size = 12, color = "black"),     
    axis.text.x = element_text(size = 12, color = "black"),   
    plot.title = element_text(size = 16, color = "black"),   
    legend.title = element_text(size = 14, color = "black"), 
    legend.text = element_text(size = 12, color = "black"), 
   strip.text = element_text(size = 14, color = "black"), 
    legend.position = "",
     legend.box = "horizontal",
    plot.background = element_rect(fill = "white"))  +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines")) +
    facet_grid(vars(), vars(site),scales = "fixed")

NH4

#nitrate
nitrogen<-read.csv("sargasso_nitrates.csv",header=T)

nitrogen$block<-as.factor(nitrogen$block)
nitrogen$trip<-as.factor(nitrogen$trip)

N <- nitrogen %>%
  filter(!(site %in% "")) %>%
  filter(!(treatment %in% "")) %>%
  group_by(trip, site, treatment) %>%
  summarize(mean_N = mean(N), N_se = sd(N) / sqrt(n())) %>%
  ggplot() +
  aes(x = trip, y = mean_N, fill = treatment) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean_N - N_se, ymax = mean_N + N_se), 
                position = position_dodge(width = 0.9), width = 0.25) +
  scale_fill_manual(values = c("lightgray", "dark gray")) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 2023")) +
  labs(
    x = " ",
    y = expression("Nitrate (mg) sediment (kg"^{-1} * ")"),
    title = "",
    fill = "Treatment") +
  theme_classic() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
   axis.title = element_text(size = 14, color = "black"),   
    axis.text = element_text(size = 12, color = "black"),     
    axis.text.x = element_text(size = 12, color = "black"),   
    plot.title = element_text(size = 16, color = "black"),   
    legend.title = element_text(size = 14, color = "black"), 
    legend.text = element_text(size = 12, color = "black"), 
   strip.text = element_text(size = 14, color = "black"), 
      strip.text.x = element_blank(), 
    legend.position = "",
    legend.box = "horizontal",
    plot.background = element_rect(fill = "white")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines")) +
  facet_grid(vars(), vars(site), scales = "fixed")
  
print(N)


#carbon
carbon<-read.csv("sargasso_nitrates.csv",header=T)

carbon$block<-as.factor(carbon$block)
carbon$trip<-as.factor(carbon$trip)

C<-carbon %>%
 filter(!(site %in% "")) %>%
 filter(!(treatment %in% "")) %>%
  group_by(trip, site, treatment) %>%
  summarize(mean_C = mean(C), C_se = sd(C) / sqrt(n())) %>%
  ggplot() +
  aes(x =trip, y = mean_C, fill = treatment) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean_C - C_se, ymax = mean_C + C_se), 
                position = position_dodge(width = 0.9), width = 0.25) +
   scale_fill_manual(values = c("lightgray", "dark gray")) +
    scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 2023") ) +
  labs(
    x = " ",
    y = expression("DOC (mg) sediment (kg"^{-1} * ")"),
    title = "",
    fill = "Treatment") +
  theme_classic()  +
   theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
   axis.title = element_text(size = 14, color = "black"),   
    axis.text = element_text(size = 12, color = "black"),     
    axis.text.x = element_text(size = 12, color = "black"),   
    plot.title = element_text(size = 16, color = "black"),   
    legend.title = element_text(size = 14, color = "black"), 
    legend.text = element_text(size = 12, color = "black"), 
   strip.text = element_text(size = 14, color = "black"), 
     strip.text.x = element_blank(), 
    legend.position = "bottom",
     legend.box = "horizontal",
    plot.background = element_rect(fill = "white"))  +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines")) +
    facet_grid(vars(), vars(site),scales = "fixed")

C


# Stack the plots vertically
stacked_plots <- NH4 / N / C

# Display the stacked plots
stacked_plots

ggsave("Figure 3.jpg", plot = stacked_plots, width = 8, height = 10, dpi = 600)

```

##CO2
```{r}
##############################gases#################################################
#gas collectors

CO2<-read.csv("gas_readings.csv",header=T)

# Recode site labels
CO2$site <- factor(CO2$site,
                   levels = c("D", "J"),
                   labels = c("Beach", "Forest"))

# Convert trip labels
CO2$trip <- factor(CO2$month,
                   levels = c("Nov", "Mar", "Aug"),
                   labels = c("Nov. 22", "Mar. 23", "Aug. 23"))

CO2graph <- ggplot(CO2, aes(x = trt, y = gCO2_m2_h)) +
  stat_summary(fun = "mean", geom = "bar", fill = "#848890") +
  stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.2) +
  theme_minimal() +
  facet_grid(vars(trip), vars(site)) +
  labs(
    x = " ",
    y = expression("Soil respiration (gCO"[2]*"/m"^2*"/h)")
  ) +
  theme_classic() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    axis.title = element_text(size = 16, color = "black"),
    axis.text = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    plot.title = element_text(size = 18, color = "black"),
    legend.title = element_text(size = 16, color = "black"),
    legend.text = element_text(size = 14, color = "black"),
    strip.text = element_text(size = 16, color = "black"),
    plot.background = element_rect(fill = "white")
  ) +
  scale_x_discrete(labels = c("Control", "Sargassum"))

CO2graph

ggsave("Figure CO2.jpg", plot = CO2graph, width = 8, height = 9, dpi = 600)



```

##interior percent cover
```{r}
#import dataset
interiorrsurvey<-read.csv("interior_survey.csv", header = TRUE)

interior_survey_long<- interiorsurvey %>%
  gather(key = "cover_type", value = "percent_cover", c(57,60))
interior_survey_long$Trip<-as.factor(interior_survey_long$Trip)


##############################interior percent cover##################################
interior_pc_cover <- ggplot(data = interior_survey_long,
            aes(
              x = Trip,
              y = percent_cover,
              shape = cover_type,
              color = Treatment
            )) +
  stat_summary(geom = "point", fun = mean, position = position_dodge(width = 1),
               size = 2L, stroke = 1) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 1),
               width = 0.2) +   
  labs(
    x = " ",
    y = "Mean percent cover",
    title = "Interior mean percent cover"
  ) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 23")
  ) +
  facet_grid(vars(), vars(Site), labeller = labeller(Site = c(Dunes = "Beach", Jungle = "Forest"))) +   
  theme_classic() +
  scale_fill_manual(values = c("grey44", "black")) +
  scale_color_manual(values = c("grey44", "black")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 16),
    axis.title = element_text(size = 16, margin = margin(t = 0, r = 10, b = 0, l = 0)),  
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 0),
    plot.title = element_text(size = 14),
    legend.title = element_text(size = 14),  
    legend.text = element_text(size = 14) ) +
  scale_shape_manual(
    values = c("circle",  "triangle"),
    labels = c("grass", "other plants")  ) +
  labs(shape = "Cover type")   
  

interior_pc_cover




#effect size graph

interioreffsize1 <-read.csv("interior_graphs.csv",header=T)
interioreffsize <-subset(interioreffsize1, Data == "eff_size")

interioreffsize<-subset(interioreffsize, !Cover == "all_plants")
interioreffsize<-subset(interioreffsize, !Cover ==  "bare" ) 
interioreffsize<-subset(interioreffsize, !Cover == "leaf_litter")
interioreffsize$Trip<-as.factor(interioreffsize$Trip)
interioreffsize$Effect_size<-as.numeric(interioreffsize$Effect_size)


#interior effect size graph
int_eff_plot <- ggplot(interioreffsize, aes(x = Trip, y = Effect_size, shape = Cover)) +
  labs(y = "Effect size", title = "") +
  geom_path(aes(group = Cover), position = position_dodge(0.5), color = "dark gray") +
  theme_classic() +
  facet_wrap(vars(Site), ncol = 2L) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23", "4" = "Aug. 23")
  ) +
    scale_shape_discrete(labels = c("grass", "other plants")) +
 theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 0),
    legend.text = element_text(size = 16),
    strip.text = element_blank() ) +
  xlab("") 

int_eff_plot

#combine 
combined_int_effgraphs <- grid.arrange(interior_pc_cover, int_eff_plot, ncol = 1)

```

##perimeter percent cover
```{r}

perimetersurvey<-read.csv("perimeter_survey.csv", header = TRUE)

perimetersurvey$Treatment<-as.factor(perimetersurvey$Treatment)
perimetersurvey$Block<-as.factor(perimetersurvey$Block)
perimetersurvey$Trip<-as.factor(perimetersurvey$Trip)

view(perimeter_short)
perimeter_short <- subset(perimeter_survey_long, !(Trip %in% c("2", "3")))

#################################perimeter percent cover#########################
perimeter_pc_cover<-ggplot(data = perimeter_short,
       aes(
         x = Trip,
         y = percent_cover,
         shape = cover_type,
         color = Treatment
       )) +
  stat_summary(geom = "point", fun = mean, position = position_dodge(width = 1),
               size = 2L, stroke = 1) +
  stat_summary(geom = "errorbar", fun.data = mean_se, position = position_dodge(width = 1),
               width = 0.2) +
  labs(
    x = " ",
    y = "Mean percent cover",
    title = "Perimeter mean percent cover"
  ) +
  scale_x_discrete(
    limits = c("1", "4"),
    labels = c("1" = "Aug. 22", "4" = "Aug. 23")
  ) +
   facet_grid(vars(), vars(Site), labeller = labeller(Site = c(Dunes = "Beach", Jungle = "Forest"))) +   
  theme_classic() +
  scale_fill_manual(values = c("grey44", "black")) +
  scale_color_manual(values = c("grey44", "black")) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 16),
    axis.title = element_text(size = 16, margin = margin(t = 0, r = 10, b = 0, l = 0)),  
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 0),
    plot.title = element_text(size = 14),
    legend.title = element_text(size = 14),  
    legend.text = element_text(size = 14) ) +
  scale_shape_manual(
    values = c("circle",  "triangle"),
    labels = c("grass", "other plants")  ) +
  labs(shape = "Cover type")   


perimeter_pc_cover


#############################################3eff sizes###############################

perimetereffsize1 <-read.csv("perimeter_graphs.csv",header=T)
perimetereffsize <-subset(perimetereffsize1, Data == "eff_size")

perimetereffsize<-subset(perimetereffsize, !Cover == "all_plants")
perimetereffsize<-subset(perimetereffsize, !Cover ==  "dead" ) 
perimetereffsize<-subset(perimetereffsize, !Cover ==  "bare" ) 
perimetereffsize$Trip<-as.factor(perimetereffsize$Trip)
perimetereffsize$Effect_size<-as.numeric(perimetereffsize$Effect_size)

perimeter_short_effsize <- subset(perimetereffsize, !(Trip %in% c("2", "3")))

per_eff_plot <- ggplot(perimeter_short_effsize, aes(x = Trip, y = Effect_size, shape = Cover)) +
  labs(y = "Effect size", title = "") +
  geom_path(aes(group = Cover), position = position_dodge(0.5), color = "dark gray") +
  theme_classic() +
  facet_wrap(vars(Site), ncol = 2L) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "4"),
    labels = c("1" = "Aug. 22", "4" = "Aug. 23")
  ) +
  scale_shape_discrete(labels = c("grass", "other plants")) +
 theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 0),
    legend.text = element_text(size = 16),
    strip.text = element_blank() ) +
  xlab("") 
per_eff_plot

#combine 
combined_perimeter_graph<- grid.arrange(perimeter_pc_cover, per_eff_plot, ncol = 1)

```

##distance effect
```{r}
dist<-read.csv("ladder.csv",header=T)
view(dist)
dist$trip<-is.facet(dist$trip)

#make data long 
laddersurvey_long <- gather(ladder, key = "cover_type", value = "percent_cover", 16:17, -c(1:5),) 

```

##arthropods
###crawling
```{r}


##################################arthropods##########################

#pitfalls
#load data
pitfalls <-read.csv("Pitfalls.csv",header=T)
pitfalls_long<- pitfalls %>%
  gather(key = "Taxa", value = "Count", c(8,13,15))
pitfalls_long$Trip<-as.factor(pitfalls_long$Trip)


# Calculate means and standard errors for each combination of Trip, Taxa, and Treatment
pitfalls_data <- pitfalls_long %>%
  group_by(Site, Trip, Treatment, Taxa) %>%
  summarize(mean_count = mean(Count), 
            std_error = sd(Count) / sqrt(n()))

pitfall_count <- ggplot(pitfalls_data) +
  aes(x = Trip, y = mean_count, shape = Taxa) + 
  geom_jitter(position = position_dodge(width = 0.5), size = 2, alpha = 0.7) +   
  geom_errorbar(aes(ymin = mean_count - std_error, ymax = mean_count + std_error),
                position = position_dodge(width = 0.5), width = 0.2) +  
  theme_minimal() +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")) +
  theme_classic() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 16),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    legend.position = ""
  ) +
  xlab("") +
  ylab("Crawling arthropod abundance") +
  facet_grid(vars(Treatment), vars(Site), scales = "fixed", 
             labeller = labeller(Treatment = c(Control = "Control", Sargassum = "Sargassum"),
                                 Site = c(Dune = "Beach", Jungle = "Forest"))) +
  theme(strip.text = element_text(size = 16)) +
  labs(title = "") +
  labs(subtitle = "") +
  labs(caption = "") +
  theme(plot.title = element_text(size = 16))

print(pitfall_count)




#####Effect sizes Graph#####
#I manually input the effect sizes, lower and upper limits to a file called pitfall_effsize.csv
pitfalleffsize <-read.csv("pitfalls_effsize.csv",header=T)

pitfalleffsize$Trip<-as.factor(pitfalleffsize$Trip)

pitfall_eff_plot <- ggplot(pitfalleffsize, aes(x = Trip, y = Effect_size, shape = Order)) +
  labs(y = "Effect size", shape = "Taxa") +
  geom_path(aes(group = Order), position = position_dodge(0.5), color = "dark gray") +
  theme_classic() +
  facet_wrap(vars(Site), ncol = 2L, labeller = as_labeller(c("Dune" = "Beach", "Jungle" = "Forest"))) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
  theme(
    strip.text = element_blank(),  # Add this line to remove facet labels on top
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    legend.position = "bottom"
  ) +
  xlab("")


pitfall_eff_plot

combined_pit_graphs <- grid.arrange(
  pitfall_count,
  pitfall_eff_plot,
  ncol = 1,
  heights = c(4, 2))

print(combined_pit_graphs)

ggsave("Figure 7.jpg", plot = combined_pit_graphs, width = 8, height = 9, dpi = 600)

```

###flying
```{r}
####################################aerial traps#############################


#load data
stickies <-read.csv("Stickies.csv",header=T)
stickies_long<- stickies %>%
  gather(key = "Taxa", value = "Count", c(9,16,15))
stickies_long$Trip<-as.factor(stickies_long$Trip)


# Calculate means and standard errors for each combination of Trip, Taxa, and Treatment
stickies_data <- stickies_long %>%
  group_by(Trip, Site, Treatment, Taxa) %>%
  summarize(mean_count = mean(Count), 
            std_error = sd(Count) / sqrt(n()))

# Create the plot with jittered dots and error bars
stickies_prc_cover <- ggplot(stickies_data) +
  aes(x = Trip, y = mean_count, shape = Taxa) + 
  geom_jitter(position = position_dodge(width = 0.5), size = 2, alpha = 0.7) +   
  geom_errorbar(aes(ymin = mean_count - std_error, ymax = mean_count + std_error),
                position = position_dodge(width = 0.5), width = 0.2) +  
  theme_classic() +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")) +
  theme_classic() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 16),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    legend.position = ""
  ) +
  xlab("") +
  ylab("Flying arthropod abundance") +
  facet_grid(vars(Treatment), vars(Site), scales = "fixed", 
             labeller = labeller(Treatment = c(Control = "Control", Sargassum = "Sargassum"),
                                 Site = c(Dune = "Beach", Jungle = "Forest"))) +
  theme(strip.text = element_text(size = 16)) +
  labs(title = "") +
  labs(subtitle = "") +
  labs(caption = "") +
  theme(plot.title = element_text(size = 16))

print(stickies_prc_cover)


#####Effect sizes Graph#####
#I manually input the effect sizes, lower and upper limits to a file called sticky_effsize.csv
stickieseffsize <-read.csv("sticky_effsize.csv",header=T)

stickieseffsize$Trip<-as.factor(stickieseffsize$Trip)


stickies_eff_plot <- ggplot(stickieseffsize, aes(x = Trip, y = Effect_size, shape = Order)) +
  labs(y = "Effect size", shape = "Taxa") +
  geom_path(aes(group = Order), position = position_dodge(0.5), color = "dark gray") +
  theme_classic() +
  facet_wrap(vars(Site), ncol = 2L, labeller = as_labeller(c("Dune" = "Beach", "Jungle" = "Forest"))) +
  geom_point(color = "black", size = 3, position = position_dodge(0.5)) +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")
  ) +
theme(
    strip.text = element_blank(),  # Add this line to remove facet labels on top
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    legend.position = "bottom"
  ) +
  xlab("")

stickies_eff_plot

# combine

combined_sticky_graphs <- grid.arrange(
  stickies_prc_cover,
  stickies_eff_plot,
  ncol = 1,
  heights = c(4, 2))

 
print(combined_sticky_graphs)

ggsave("Figure 8.jpg", plot = combined_sticky_graphs, width = 8, height = 9, dpi = 600)

```

#Supplementary graphs
##pile vol

```{r}
####################may need to cut this graph for publication due to too many figures
#####################################pile volumes#####################################

piles <-read.csv("Sargassum_piles.csv",header=T)
piles2<-subset(piles, Treatment== "Sargassum")
piles2$Trip<-as.factor(piles2$Trip)
piles2$Block<-as.factor(piles2$Block)
# Calculate the means by month and location

mean_data <- piles2 %>%
  group_by(Trip, Location) %>%
  summarize(mean_ellipsoid_vol = mean(Ellipsoid_prc_D))


# Create the scatter plot with mean points connected by a line
scatterplot <- ggplot(data = piles2, aes(x = Trip, y = Ellipsoid_prc_D, fill = Location)) +
  geom_point(size = 2, shape = 21, aes(fill = Location)) +
  geom_line(data = mean_data, aes(x = Trip, y = mean_ellipsoid_vol, group = Location), color = "gray")

# Customize the plot appearance
vol_x_time <- scatterplot +
  labs(x = "", y = "Pile volume (%)") +
  scale_color_manual(values = c("black", "white")) +
  scale_fill_manual(values = c("black", "white")) +
  scale_x_discrete(
    limits = c("1", "2", "3", "4"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar.23", "4" = "Aug.23")
  ) +
  theme_classic() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    axis.title = element_text(size = 16, color = "black"),    
    axis.text = element_text(size = 14, color = "black"),     
    axis.text.x = element_text(size = 14, color = "black"),   
    plot.title = element_text(size = 18, color = "black"),    
    legend.title = element_text(size = 16, color = "black"),  
    legend.text = element_text(size = 14, color = "black")  )

vol_x_time
```

## full ladder survey
```{r}
ladder_survey_full <-read.csv("ladder.csv",header=T)  

ladder_survey_long<- ladder_survey_full %>%
  gather(key = "cover_type", value = "percent_cover", c(10,11))
ladder_survey_long$Trip<-as.factor(ladder_survey_long$Trip)
ladder_survey_long$Quadrat<-as.factor(ladder_survey_long$Quadrat)

#remove zero points since there is overlap witht he actual plot
ladder_survey_short2 <- subset(ladder_survey_long, !Quadrat == "0")

#sumarize data for the graph
ladder_survey_summed <- ladder_survey_short2 %>%
  group_by(Site, Trip, Dist_from_edge, Treatment, cover_type) %>%
  summarize(mean_cover = mean(percent_cover),
            mean_se = sd(percent_cover) / sqrt(n()))

#subset grass to make graph
ladder_survey_grass<-subset(ladder_survey_summed, cover_type == "Bermuda")

#grass graph
ladder_sargasso_graph <- ggplot(ladder_survey_grass) +
  aes(x = Trip, y = mean_cover, fill = Treatment) +  
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_errorbar(aes(ymin = mean_cover - mean_se, ymax = mean_cover + mean_se), 
                position = position_dodge(width = 0.7), width = 0.2) +
  scale_fill_manual(values = c("lightgray", "darkgray"),
                    labels = c("Control", "Sargassum")) +
  labs(
    x = "",
    y = "Grass percent cover",
    fill = "Treatment",  ) +

  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")) +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14, color = "black"),
    axis.title = element_text(size = 16, color = "black"),
    axis.text = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    plot.title = element_text(size = 12, color = "black"),
    legend.title = element_text(size = 16, color = "black"),
    legend.text = element_text(size = 14, color = "black"),
    strip.text = element_text(size = 14, color = "black")) +
  facet_grid(vars(Dist_from_edge), vars(Site))

print(ladder_sargasso_graph)

#now lets repeat this for other plants

#subset other plants for second graph
ladder_survey_others<-subset(ladder_survey_summed, cover_type == "Other_pants")

#graph it

ladder_sargasso_graph2 <- ggplot(ladder_survey_others) +
  aes(x = Trip, y = mean_cover, fill = Treatment) +  
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  geom_errorbar(aes(ymin = mean_cover - mean_se, ymax = mean_cover + mean_se), 
                position = position_dodge(width = 0.7), width = 0.2) +
  scale_fill_manual(values = c("lightgray", "darkgray"),
                    labels = c("Control", "Sargassum")) +
  labs(
    x = "",
    y = "Other plant percent cover",
    fill = "Treatment",  ) +

  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")) +
  theme_minimal() +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14, color = "black"),
    axis.title = element_text(size = 16, color = "black"),
    axis.text = element_text(size = 14, color = "black"),
    axis.text.x = element_text(size = 14, color = "black"),
    plot.title = element_text(size = 12, color = "black"),
    legend.title = element_text(size = 16, color = "black"),
    legend.text = element_text(size = 14, color = "black"),
    strip.text = element_text(size = 14, color = "black")) +
  facet_grid(vars(Dist_from_edge), vars(Site))


print(ladder_sargasso_graph2)

#combine graphs
combined_ladder_graphs <- grid.arrange(
  ladder_sargasso_graph,
  ladder_sargasso_graph2,
  ncol = 1,
  heights = c(4, 4))

print(combined_ladder_graphs)
```

##full perimeter percent cover
```{r}
#import dataset
perimetersurvey_full<-read.csv("perimeter_survey.csv", header = TRUE)

perimeter_prc_long<-perimetersurvey_full %>%
  gather(key = "cover_type", value = "prc_cover", c(7:58))

perimeter_prc_long$Trip<-as.factor(perimeter_prc_long$Trip) 
perimeter_prc_long$Block<-as.factor(perimeter_prc_long$Block) 

#select only relevant columns
perimeter_prc_long <- perimeter_prc_long %>% select( c(1:5,15,16))
#remove zeros so the data is not so cluttered
# Remove rows where prc_cover is zero
perimeter_prc_long <- subset(perimeter_prc_long, prc_cover != 0)



perimeter_prc_long %>%
  filter(Site %in% "Jungle") %>%
  ggplot() +
  aes(x = Trip, fill = cover_type, weight = prc_cover) +
  geom_bar(position = "fill") +
  theme_minimal() +
  facet_grid(vars(Block), vars(Treatment)) +
  labs(y = "Proportion")


perimeter_prc_long %>%
 filter(Site %in% "Dunes") %>%
 ggplot() +
 aes(x = Trip, fill = cover_type, weight = prc_cover) +
 geom_bar(position = "fill") +
 theme_minimal() +
 facet_grid(vars(Block), vars(Treatment)) +
  labs(y = "Proportion")

```

##full interior percent cover
```{r}
interiorsurvey_full<-read.csv("interior_survey.csv", header = TRUE)

#make data long with species types all stacked
interior_prc_long<-interiorsurvey_full %>%
  gather(key = "cover_type", value = "prc_cover", c(6:55))

interior_prc_long$Trip<-as.factor(interior_prc_long$Trip) 
interior_prc_long$Block<-as.factor(interior_prc_long$Block) 
#remove zeros so the data is not so cluttered
# Remove rows where prc_cover is zero
interior_prc_long <- subset(interior_prc_long, prc_cover != 0)
#select only relevant columns
interior_prc_long <- interior_prc_long %>% select( c(1:5,14,15))
  

interior_prc_long %>%
 filter(Site %in% "Dunes") %>%
 ggplot() +
  aes(x = Trip, fill = cover_type, weight = prc_cover) +
  geom_bar(position = "fill") +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_grid(vars(Block), vars(Treatment))+
  labs(y = "Proportion")

interior_prc_long %>%
 filter(Site %in% "Jungle") %>%
 ggplot() +
  aes(x = Trip, fill = cover_type, weight = prc_cover) +
  geom_bar(position = "fill") +
  scale_fill_hue(direction = 1) +
  theme_minimal() +
  facet_grid(vars(Block), vars(Treatment)) +
   labs(y = "Proportion")

```



##practice
```{r}


# Create the plot with jittered dots and error bars
pitfall_count <- ggplot(pitfalls_data) +
  aes(x = Trip, y = mean_count, shape = Taxa) +  # shape by Site
  geom_jitter(position = position_dodge(width = 0.5), size = 2, alpha = 0.7) +   
  geom_errorbar(aes(ymin = mean_count - std_error, ymax = mean_count + std_error),
                position = position_dodge(width = 0.5), width = 0.2) +  
  theme_minimal() +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")) +
theme_minimal() +
theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_blank(),
      legend.position = "bottom") +
  xlab("") +
  ylab("Crawling arthropod abundance") +
  
  facet_grid(vars(Treatment), vars(Site), scales = "fixed", space = "fixed", labeller = labeller(Treatment = c(Control = "control", Sargassum = "sargasso"))) +
    guides(shape = FALSE)  


pitfall_count


# combine

combined_pitfall_graphs <- grid.arrange(
  pitfall_count,
  pitfall_eff_plot,
  ncol = 1,
  heights = c(3, 2))

# Print or display the combined plot
print(combined_pitfall_graphs)



# Create the plot with jittered dots and error bars
stickies_prc_cover <-ggplot(stickies_data) +
  aes(x = Trip, y = mean_count, shape = Taxa) +  
  geom_jitter(position = position_dodge(width = 0.5), size = 2, alpha = 0.7) +   
  geom_errorbar(aes(ymin = mean_count - std_error, ymax = mean_count + std_error),
                position = position_dodge(width = 0.5), width = 0.2) +   
  theme_minimal() +
  scale_x_discrete(
    limits = c("1", "2", "3"),
    labels = c("1" = "Aug. 22", "2" = "Nov. 22", "3" = "Mar. 23")) +
theme_minimal() +
theme(
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    panel.spacing = unit(0.2, "lines"),
    text = element_text(size = 14),
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
    axis.text.x = element_text(size = 14),
    plot.title = element_text(size = 12),
    legend.title = element_text(size = 16),
    legend.text = element_text(size = 14),
    strip.text = element_blank()
  ) +
  xlab("") +
  ylab("Flying arthropod abundance") +
  facet_grid(vars(Treatment), vars(Site), scales = "free", space = "fixed", labeller = labeller(Treatment = c(Control = "control", Sargassum = "sargasso"))) +
    guides(shape = FALSE)   
stickies_prc_cover
































sargman<-read.csv("nmec.csv",header=T)

# Assuming your_data is your original data frame
sargman2 <- sargman %>%
  pivot_longer(cols = c(top_layer, bottom_layer), names_to = "layer", values_to = "species")










# Assuming your_data is your data frame
wide_data <- reshaped_data %>%
  pivot_wider(names_from = species,
              values_from = species,
              values_fn = length,
              values_fill = 0,
              names_prefix = "")

# Print the resulting wide_data
view(wide_data)


long_data <- gather(wide_data, key = "species", value = "count", 14:81, -c(1:13))


# Group the data by quadrat_code and survey_date, then calculate the sum of other columns
collapsed_df <- long_data %>%
  group_by(quadrat_code, survey_date) %>%
  summarize(total = sum(count),   # Replace 'column1' with the column you want to sum
            # Add more columns you want to sum here
            )

# Replace 'column1', etc. with the actual column names you want to sum.


# Group the data by 'quadrat_code' and 'survey_date', and summarize the data
summary_df <- wide_data %>%
  group_by(quadrat_code, survey_date) %>%
  summarize(total = sum(column_to_sum),  # Replace 'column_to_sum' with the column you want to summarize
            mean_value = mean(another_column))  # Replace 'another_column' with another column you want to summarize

# Replace 'column_to_sum' and 'another_column' with the actual columns you want to summarize.



# Assuming your data frame is named 'df'
library(dplyr)

# Add a new column 'total' that contains the sum of columns 3 through 30
wide_data2 <- wide_data %>%
  mutate(total = rowSums(across(14:81)))

# This calculates the sum of columns 3 through 30 for each row.

view(wide_data2)

```

